### 2021-04-14 analysis/index.Rmd
wflow_publish(c("analysis/plot.Rmd", "analysis/_site.yml", "analysis/index.Rmd"),     republish = TRUE)

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

This website collects comparisons of Amphibolis and Posidonia comparative analyses.

[Repeat comparison plots](plot.html)

[Phylogenies](OTT.html)