### 2021-04-14 analysis/plot.Rmd

---
title: "Untitled"
author: "Philipp Bayer"
date: "4 March 2021"
output: html_document
---

```{r setup}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(RColorBrewer)
library(patchwork)
```

```{r}
# reps <- data.frame(
#   stringsAsFactors = FALSE,
#              Class = c("LINEs","LTRs","DNA",
#                        "Unclassified","Non-repetitive","LINEs","LTRs","DNA","Unclassified",
#                        "Non-repetitive","LINEs","LTRs","DNA","Unclassified",
#                        "Non-repetitive","LINEs","LTRs","DNA","Unclassified","Non-repetitive"),
#           Assembly = c("A. antarctica","A. antarctica",
#                        "A. antarctica","A. antarctica","A. antarctica","P. australis",
#                        "P. australis","P. australis","P. australis","P. australis","Z. marina",
#                        "Z. marina","Z. marina","Z. marina","Z. marina","Z. muelleri",
#                        "Z. muelleri","Z. muelleri","Z. muelleri","Z. muelleri"),
#                 BP = c(805898L,45776664L,37822890L,
#                        1546483L,158725395L,38319733L,611240734L,100034681L,
#                        37142931L,428369702L,3103364L,123623579L,27390818L,
#                        6060480L,96439714L,14331182L,164253396L,142828758L,
#                        21922574L,274954485L),
#            Percent = c(0.33,18.71,15.46,0.63,
#                        64.87131,3.15,50.3,8.23,3.06,35.25364,1.19,47.46,10.52,
#                        2.33,37.02,2.32,26.57,23.1,3.55,44.46)
# )

reps <- readxl::read_xlsx('data/Repeat_coding.xlsx')

reps <- reps %>% dplyr::filter(Class != 'Total')
reps$Class <- factor(reps$Class, levels = c('LINEs', 'LTRs', 'DNA', 'Unclassified', 'Total CDS', 'Non-repetitive'))
reps
```

```{r}
p1 <- reps %>% ggplot(aes(x=Assembly, fill=Class, y = BP/1000000)) + 
geom_bar(position='stack', stat='identity') + ylab('Size (Mbp)') +
  scale_fill_brewer(palette='Dark2') +
  theme(axis.text.x =  element_text(face="italic"))

p1
```

```{r}
p2 <- reps %>% ggplot(aes(x=Assembly, fill=Class, y = Percent)) + geom_bar(position='stack', stat='identity')+
  scale_fill_brewer(palette='Dark2') +
  theme(axis.text.x =  element_text(face="italic")) +
  ylab('Percent\n of assembly')
p2
```
```{r fig.width = 10, dpi=300}
p1/p2 + plot_annotation(tag_levels = 'A')
```

```{r}
reps %>% filter(Class=='Non-repetitive') %>% ggplot(aes(x=Assembly, fill=Class, y = BP)) + geom_bar(position='stack', stat='identity')

```