### 2021-10-07 analysis/GOenrichment.Rmd
wflow_publish(files = c("analysis/*"))

---
title: "GO enrichment"
author: "Philipp Bayer"
date: "01/10/2021"
output: html_document
---

```{r setup, include=FALSE}
library(topGO)
library(tidyverse)
library(rvest)
library(wesanderson)
library(httr)
library(stringi)
library(ggrepel)
library(eulerr)
library(UpSetR)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Lost in seagrasses

The GO enrichment does not work well on my laptop so I'm setting this to eval=FALSE and run it on a remote server.
The script which writes the input files is a Python script in the scripts/ folder: findClustersUniqueToAquatic.py. This script parses Orthofinder output to pull out genome-specific groups and genes.

```{r eval=FALSE}
# give properly formatted background in format: GO:0005838	GSBRNA2T00088508001;GSBRNA2T00088313001;GSBRNA2T00035842001 
#annAT <- readMappings('BACKGROUND.txt', sep="\t", IDsep=";")
#save(annAT, file='annAtObject.RData')

load('annAtObject.RData')
allgenes <- unique(unlist(annAT))

compare <- function(genelistfile, outname, allgenes, annAT) {
  # give file with your genes of interest, one gene_id per line

  mygenes <-scan(genelistfile ,what="")
  geneList <- factor(as.integer(allgenes %in% mygenes))
  names(geneList) <- allgenes

  GOdata <-new ("topGOdata", ontology = 'BP', allGenes = geneList, nodeSize = 5, annot=annFUN.GO2genes, GO2genes=annAT)
  # using ClassicCount:
  #test.stat <-new ("classicCount", testStatistic = GOFisherTest, name = "Fisher Test")
  #resultsFisherC <-getSigGroups (GOdata, test.stat)
  # using weight01:
  weight01.fisher <- runTest(GOdata, statistic = "fisher")
  # using ClassicCount:
  # allRes <- GenTable(GOdata, classicFisher= resultsFisherC, topNodes = 30)
  # using weight01:
  allRes <- GenTable(GOdata, classicFisher=weight01.fisher,topNodes=30)#,topNodes=100)
  names(allRes)[length(allRes)] <- "p.value"
  p_values <- score(weight01.fisher)
  adjusted_p <- p.adjust(p_values)
  adjusted_p[adjusted_p < 0.05] %>% enframe() %>% write_csv('data/' + outname)

}

```

# lost in seagrasses, each one uniquely vs all terrestrials
```{r eval=FALSE}
compare('lost_in_amphi_vs_all.txt', 'missing_amphi_vs_all_GO.txt', allgenes, annAT)
compare('lost_in_posi_vs_all.txt', 'missing_posi_vs_all_GO.txt', allgenes, annAT)
compare('lost_in_zmar_vs_all.txt', 'missing_zmar_vs_all_GO.txt', allgenes, annAT)
compare('lost_in_zmuel_vs_all.txt', 'missing_zmuel_vs_all_GO.txt', allgenes, annAT)
```

# lost in seagrasses vs all terrestrials
Here we compare GO terms for seagrasses and aquatics (seagrasses+duckweeds) vs all terrestrials

```{r eval=FALSE}

compare('lost_in_seagrasses.txt', 'missing_seagrasses_GO.txt', allgenes, annAT)
compare('lost_in_aquatics.txt', 'missing_aquatics_GO.txt', allgenes, annAT)
compare('only_in_seagrasses.txt', 'only_seagrasses_GO.txt', allgenes, annAT)

```

# Seagrass only comparisons

Now we compare seagrasses within each other.

For the seagrass-only comparisons, I'm using a Seagrass-only background as that makes more biological sense to me.s

```{r eval=FALSE}
# give properly formatted background in format: GO:0005838	GSBRNA2T00088508001;GSBRNA2T00088313001;GSBRNA2T00035842001 
#sannAT <- readMappings('SEAGRASSBACKGROUND.txt', sep="\t", IDsep=";")
#save(sannAT, file='sannAtObject.RData')

load('annAtObject.RData')
sallgenes <- unique(unlist(sannAT))

compare('lost_in_amphi.txt', 'lost_in_amphi_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('lost_in_posi.txt', 'lost_in_posi_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('lost_in_zmar.txt', 'lost_in_zmar_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('lost_in_zmuel.txt', 'lost_in_zmuel_vs_seagrasses_GO.txt', sallgenes, sannAT)
```

# Present in seagrasses

The opposite - which GO-terms are present only in one of the four species?

```{r eval=FALSE}
compare('only_in_amphi.txt', 'only_in_amphi_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('only_in_posi.txt', 'only_in_posi_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('only_in_zmar.txt', 'only_in_zmar_vs_seagrasses_GO.txt', sallgenes, sannAT)
compare('only_in_zmuel.txt', 'only_in_zmuel_vs_seagrasses_GO.txt', sallgenes, sannAT)
```

Alright now we have all these different GO terms in all these files - we can send them to revigo for visualiation and some deduplication!

# Revigo

This code is based on http://revigo.irb.hr/CodeExamples/revigo.R.txt

```{r}
results_list <- list()

for (f in list.files('./data/', pattern='GO.txt')){
  filename <- paste('./data/', f, sep='')
  go_and_pvalues <- readChar(filename, file.info(filename)$size)
  go_and_pvalues <- gsub(',', ' ', go_and_pvalues)

  httr::POST(
    url = "http://revigo.irb.hr/StartJob.aspx",
    body = list(
      cutoff = "0.7",
      valueType = "pvalue",
     # speciesTaxon = "4577", # zea mays
     #speciesTaxon = '39947', # japonica
     speciesTaxon = '3702', # arabidopsis
      measure = "SIMREL",
      goList = go_and_pvalues
    ),
    # application/x-www-form-urlencoded
    encode = "form"
  ) -> res

  dat <- httr::content(res, encoding = "UTF-8")

  jobid <- jsonlite::fromJSON(dat,bigint_as_char=TRUE)$jobid

  # Check job status
  running <- "1"
  while (running != "0" ) {
      httr::POST(
        url = "http://revigo.irb.hr/QueryJobStatus.aspx",
        query = list( jobid = jobid )
      ) -> res2
      dat2 <- httr::content(res2, encoding = "UTF-8")
      running <- jsonlite::fromJSON(dat2)$running
      Sys.sleep(1)
  }

  # Fetch results
  httr::POST(
    url = "http://revigo.irb.hr/ExportJob.aspx",
    query = list(
      jobid = jobid, 
      namespace = "1",
      type = "CSVTable"
    )
  ) -> res3

  dat3 <- httr::content(res3, encoding = "UTF-8")

  dat3 <- stri_replace_all_fixed(dat3, "\r", "")
  # Now we have a csv table in a string!

  # read_csv does not like the ', ', it wants ','
  dat <- read_csv(gsub(', ', ',', dat3), show_col_types = FALSE)

  # do we even have results?
  if(nrow(dat) == 0){next}
  results_list[[f]] <- dat

}
```

OK we have a list with all results in a big list. Now we can plot!

```{r warning=FALSE, echo=FALSE, message=FALSE}
# lots of warnings and messages here so I'm hiding these
plot_list <- list()
for (f in names(results_list)) {
  dat <- results_list[[f]]
  if(nrow(dat) == 0) {next}

  names(dat) <- c("term_ID","description","frequency", 'plot_X', 'plot_Y', 'log_size', 'value', 'uniqueness', 'dispensability', 'eliminated', 'representative')
  one.data <- dat

  one.data <- one.data [(one.data$plot_X != "null" & one.data$plot_Y != "null") & (! is.na(one.data$frequency)) & (one.data$value != 'null'), ];
  one.data <- type_convert(one.data)
  one.data$pvalue <- p.adjust(exp(one.data$value))
  #one.data <- one.data %>% filter(pvalue < 0.05) # NO FILTERING OF P-VALUES - for now?

  p1 <-
    ggplot(data = one.data, aes(plot_X, plot_Y, label = description)) +
    geom_point(aes(colour = log(pvalue), size = log_size), alpha = I(0.6)) +
    scale_size_area() +
    scale_colour_gradientn(colours = rev(wes_palette("Zissou1", 100, type = "continuous")), 
                           limits = c(min(log(one.data$pvalue)), 0)) +
    geom_point(
      aes(plot_X, plot_Y, size = log_size),
      shape = 21,
      fill = "transparent",
      colour = I (alpha ("black", 0.6))
    ) +
    scale_size_area() +
    geom_label_repel(max.overlaps = 5, box.padding = 0.5,
                     aes(point.size = log_size)) +
    theme_minimal() +
    labs(y = "Semantic space x",
         x = "Semantic space y",
         color = 'Log(p-value)',
         size = 'Frequency') +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    ) +
    ggtitle(f)

  plot_list[[f]] <- p1

}
```

# Venn diagram GO overlaps

Let's also check which GO terms overlap between the 4 seagrasses!

Let's have a look at all of these plots. Manually zooming in leads to ggrepel reloading labels, so on the small scale a lot of these plots don't have labels.
```{r}
plot_list
```

How many shared lost GO-terms are there? Hopefully, all four species will have lost the most GO-terms.
```{r}
a <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$Name,
          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name,
          `A. thaliana` = results_list$missing_arabidopsis_vs_all_GO.txt$Name)

a_go_ids <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$TermID,
          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$TermID,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$TermID,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$TermID,
          `A. thaliana` = results_list$missing_arabidopsis_vs_all_GO.txt$TermID)
```

```{r}
plot(euler(a), 
     quantities = TRUE,
     fill = rev(wes_palette("Zissou1", 15, type = 'continuous')),
    alpha = 0.5,
     labels = list(font = 4))
```

```{r}
upset(fromList(a), order.by='freq', )
```

```{r}
a_no_ara <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$Name,
          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name)
```

What are the shared GO-terms in seagrasses, WITHOUT the Ara loss??
```{r}
setdiff(Reduce(intersect, a_no_ara), Reduce(intersect, a)) %>% enframe() %>% writexl::write_xlsx('./data/shared_lost_genes.xlsx')
```

What if we remove Posidonia?

```{r}
b <- list(`A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name)
intersections <- Reduce(intersect, b)
intersections[grepl('ethylene', intersections)]
```
OK we need a big list of all GO-terms here - which GO-term is lost in which species. That will be a supplementary table.

```{r}
all_species <- c("P. australis","A. antarctica","Z. marina","Z. muelleri", 'A. thaliana')  
all_go_terms <- Reduce(union, a)
all_go_ids <- Reduce(union, a_go_ids)
results_d <- data.frame('GOID' = character(),
                        'GO' = character(),
                        'P. australis' = character(),
                        'A. antarctica' = character(),
                        'Z. marina' = character(),
                        'Z. muelleri' = character(),
                        'A. thaliana' = character())

for (index in seq_along(all_go_terms)){
  go <- all_go_terms[index]
  go_id <- all_go_ids[index]
  specs <- c()
  for (species in names(a)) {
    if ( length(a[[species]][grep(paste('^', go, '$', sep=''), a[[species]])]) > 0 ) {
      specs <- c(specs, species)
    }
  }

  results_d[index,] <- c(go_id, go, gsub('FALSE', 'Present', gsub('TRUE', 'Lost', all_species %in% specs)))
}
```

```{r}
writexl::write_xlsx(results_d, 'data/Lost_GO_terms_in_five_species.xlsx')
```

# after filtering for plant-specific GO-terms

We will use the GO-terms that are plant-specific as identified by the GOMAP paper.
See https://github.com/wkpalan/GOMAP-maize-analysis/blob/main/6.plantSpecific/1.getSppSpecific.R
or https://plantmethods.biomedcentral.com/articles/10.1186/s13007-021-00754-1

```{r}
go_plant <- read_tsv('https://raw.githubusercontent.com/wkpalan/GOMAP-maize-analysis/main/data/go/speciesSpecificGOTerms.txt')
```

```{r}
# taxon 33090 is Viridiplantae
plantSpecificGO <- go_plant %>% dplyr::filter(`NCBITaxon:33090`==1) %>% pull(GOterm)
plantSpecificGO <- c(plantSpecificGO,c("GO:0005575","GO:0008150","GO:0003674"))
```

```{r}
results_d %>% filter(GOID %in% plantSpecificGO) %>% writexl::write_xlsx('data/Lost_GO_terms_in_five_species.PlantSpecific.xlsx')
```

Now let's redo the Venn diagram with those filtered GO-terms

```{r}
filters <- lapply(a_go_ids, function(ch) ch %in% plantSpecificGO)

newa <- list()
for (species in names(filters)) {
  before <- a[[species]]
  after <- before[filters[[species]]]
  newa[[species]] <- after
}

```

```{r}
plot(euler(newa), 
     quantities = TRUE,
     fill = rev(wes_palette("Zissou1", 15, type = 'continuous')),
    alpha = 0.5,
     labels = list(font = 4))
```

Not much difference?