### 2021-10-07 analysis/Orthofinder.Rmd

---
title: "Clustering/Orthofinder analysis"
author: "Philipp Bayer"
date: "15 April 2021"
output: html_document
---

```{r setup}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(RColorBrewer)
library(patchwork)
library(UpSetR)

```

```{r}
groups <- read_tsv('./data/Orthogroups.tsv')

names(groups) <- c('Orthogroup', 'A. antarctica', 'A. trichopada', 'B. distachyon', 'C. reinhardtii', 'L. gibba', 'O. sativa', 'P. australis', 'P. patens', 'P. trichocarpa', 'S. moellendorffii', 'S. polyrhiza', 'A. thaliana', 'T. parvula', 'V. vinifera', 'Z. mays', 'Z. muelleri', 'Z. marina', 'O. lucimarinus', 'W. australis')
```

```{r}
# for upsetr, we need to know only which OG-groups are shared between species, the actual genes don't matter
per_spec <- groups %>% pivot_longer(-Orthogroup) %>% 
  filter(!is.na(value)) %>% # species not in an orthogroup are still listed, they just have NA genes for this group
  select(-value) # don't need all gene names, speed things up
```

```{r}
# now I want the data in this format:
# listInput <- list(one = c(1, 2, 3, 5, 7, 8, 11, 12, 13), two = c(1, 2, 4, 5, 
#   10), three = c(1, 5, 6, 7, 8, 9, 10, 12, 13))
x <- per_spec %>% 
  select(name, Orthogroup) %>% # turn the table around
  deframe() # convert to named vector
```

```{r}
mylist <- lapply(split(x, names(x)), unname) # yuck - ugly code to convert the named vector to a list
```

```{r}
x <- upset(fromList(mylist), order.by='freq', nsets = length(groups) - 1)
```

```{r fig.width=10}
x
```

Let's get the species-only cluster numbers

```{r}
species_specific_orthos <- per_spec %>% 
  group_by(Orthogroup) %>% 
  summarise(counts = length(name)) %>%
  filter(counts == 1)
```

```{r}
per_spec %>% 
  filter(Orthogroup %in% species_specific_orthos$Orthogroup) %>% 
  group_by(name) %>% 
  count() %>% 
  arrange(n) %>% 
  knitr::kable()
```

How many orthogroups are shared between the four seagrasses?

```{r}
newlist <- mylist[c('A. antarctica', 'Z. marina', 'P. australis', 'Z. muelleri')]

x <- upset(fromList(newlist), order.by='freq', nsets = 4)

```
```{r}
x
```