### 2021-10-07 analysis/index.Rmd

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

This website collects comparisons of *Amphibolis* and *Posidonia* comparative analyses.

[Repeat comparison plots](plot.html)

[Phylogenies](OTT.html)

[Clustering results](Orthofinder.html)

[R gene plotting](plotRgenes.html)

[Gene count association](associateGenes.html)

[GO enrichment](GOenrichment.html)