### 2021-10-07 analysis/plotGenes.Rmd

---
title: "Making gene number plots"
author: "Philipp Bayer"
date: "14/09/2021"
output: html_document
---
```{r}
library(tidyverse)
```

This page plots numbers of genes based on GO-terms, no clustering involved.
```{r}
df <- data.frame(
  stringsAsFactors = FALSE,
           Species = c("A. antarctica","O. sativa",
                       "P. australis","Z. muelleri","Z. marina"),
          Salinity = c(3L, 35L, 3L, 9L, 2L),
               PPR = c(254L, 590L, 291L, 913L, 528L),
        Heat.shock = c(42L, 102L, 56L, 139L, 46L),
    Pectinesterase = c(21L, 51L, 30L, 76L, 67L)
)
```

```{r}
df %>% pivot_longer(-Species) %>% ggplot(aes(x=factor(Species, levels=rev(c('O. sativa', 'A. antarctica', 'P. australis', 'Z. marina', 'Z. muelleri'))), y=value)) + 
  geom_bar(stat='identity') + 
  facet_wrap(~name, scales='free_x' ) +
  ylab('Count') + xlab('Species') + 
  cowplot::theme_minimal_hgrid() + coord_flip() +
  theme(axis.text.y = element_text(face = "italic"))
```