### 2021-10-07 analysis/plotRgenes.Rmd

---
title: "R genes"
output: html_notebook
---
```{r setup}
library(tidyverse)
library(wesanderson)
```

Visualising R-gene differences here

```{r}
pal <- wes_palette("Zissou1", 3, type = "continuous")

df <- readxl::read_xlsx('./data/R_genes.xlsx')
df %>% ggplot(aes(fill=Class, x=Genome, y=Count)) + geom_bar(stat='identity') +
    scale_fill_manual(values=pal) +  
 coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic"))

```

```{r}
pal <- wes_palette("Zissou1", 4, type = "continuous")

df2 <- df %>% mutate(Class2= case_when(str_detect(Subclass, pattern = 'TN') ~ 'NLR (TNL)',
                                str_detect(Subclass, pattern = '^R') ~ Subclass,
                                TRUE ~ 'NLR (CNL)'))
df2 %>% ggplot(aes(fill=Class2, x=Genome, y=Count)) + geom_bar(stat='identity') +
    scale_fill_manual(values=pal)  + coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic")) 
```

```{r}
df2 %>% dplyr::filter(Genome %in% c('P. australis', 'Z. marina', 'Z. muelleri', 'O. sativa', 'A. antarctica', 'A. thaliana')) %>% 
  ggplot(aes(fill=Class2, x=factor(Genome, levels=c('Z. marina', 'Z. muelleri', 'P. australis', 'A. antarctica', 'O. sativa', 'A. thaliana' )), y=Count)) + 
  geom_bar(stat='identity') +
    scale_fill_manual(values=pal)  + coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic")) +
  xlab('Genome') +
  labs(fill='Class')
```