### 2021-10-27 analysis/Orthofinder.Rmd
Add missing files

---
title: "Clustering/Orthofinder analysis"
author: "Philipp Bayer"
date: "15 April 2021"
output: html_document
---

```{r setup}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(RColorBrewer)
library(patchwork)
library(UpSetR)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

```{r}
groups <- read_tsv('./data/Orthogroups.tsv')

names(groups) <- c('Orthogroup', 'A. antarctica', 'A. trichopada', 'B. distachyon', 'C. reinhardtii', 'L. gibba', 'O. sativa', 'P. australis', 'P. patens', 'P. trichocarpa', 'S. moellendorffii', 'S. polyrhiza', 'A. thaliana', 'T. parvula', 'V. vinifera', 'Z. mays', 'Z. muelleri', 'Z. marina', 'O. lucimarinus', 'W. australis')
```

```{r}
# for upsetr, we need to know only which OG-groups are shared between species, the actual genes don't matter
per_spec <- groups %>% pivot_longer(-Orthogroup) %>% 
  filter(!is.na(value)) %>% # species not in an orthogroup are still listed, they just have NA genes for this group
  select(-value) # don't need all gene names, speed things up
```

```{r}
# now I want the data in this format:
# listInput <- list(one = c(1, 2, 3, 5, 7, 8, 11, 12, 13), two = c(1, 2, 4, 5, 
#   10), three = c(1, 5, 6, 7, 8, 9, 10, 12, 13))
x <- per_spec %>% 
  select(name, Orthogroup) %>% # turn the table around
  deframe() # convert to named vector
```

```{r}
mylist <- lapply(split(x, names(x)), unname) # yuck - ugly code to convert the named vector to a list
```

```{r}
x <- upset(fromList(mylist), order.by='freq', nsets = length(groups) - 1)
```

```{r fig.width=10}
x
```

Let's get the species-only cluster numbers

```{r}
species_specific_orthos <- per_spec %>% 
  group_by(Orthogroup) %>% 
  summarise(counts = length(name)) %>%
  filter(counts == 1)
```

```{r}
per_spec %>% 
  filter(Orthogroup %in% species_specific_orthos$Orthogroup) %>% 
  group_by(name) %>% 
  count() %>% 
  arrange(n) %>% 
  knitr::kable()
```

How many orthogroups are shared between the four seagrasses?

```{r}
newlist <- mylist[c('A. antarctica', 'Z. marina', 'P. australis', 'Z. muelleri')]

x <- upset(fromList(newlist), order.by='freq', nsets = 4)

```
```{r}
x
```

# Connect Orthofinder results with functional table of Arabidopsis genes

File comes from https://www.arabidopsis.org/download_files/Genes/TAIR10_genome_release/TAIR10_functional_descriptions

```{r eval=FALSE}
download.file('https://www.arabidopsis.org/download_files/Genes/TAIR10_genome_release/TAIR10_functional_descriptions', 'data/TAIR10_functional_descriptions')
```
```{r}
functions <- read_tsv('data/TAIR10_functional_descriptions')
```
I also downloaded the gene symbols:

```{r eval=FALSE}
esearch -db gene -query "Arabidopsis thaliana [ORGN]" | esummary | xtract -pattern DocumentSummary -element Name,OtherAliases |  awk -F "\t|," '{OFS="\t"}{print $2,$1}' > arabidopsis_gene_symbols.txt
```

```{r}
symbols <- read_tsv('./data/arabidopsis_gene_symbols.txt', col_names = c('Gene','Symbol'))
```

OK now I need to flip the Orthofinder table to then join the functional table

```{r}
ara_groups <- groups %>% select(Orthogroup, `A. thaliana`) %>% filter(!is.na(`A. thaliana`)) %>% separate_rows(Orthogroup, `A. thaliana`, convert=T)
ara_groups <- ara_groups %>% separate(`A. thaliana`, c('Gene','Number'), remove=FALSE)
ara_groups <- left_join(ara_groups, symbols)
ara_groups$Number <- NULL
head(ara_groups)
```

```{r}
ara_joined <- left_join(ara_groups, functions, by=c(`A. thaliana`='Model_name'))
```

You know what? we shouldn't directly compare a dicot (A. thaliana) with monocots (seagrasses), should also add another regular monocot (rice)

```{r}
posi_groups <- groups %>% select(Orthogroup, `P. australis`) %>% filter(!is.na(`P. australis`)) %>% separate_rows(Orthogroup, `P. australis`, convert=T, sep = ', ')
amphi_groups <- groups %>% select(Orthogroup, `A. antarctica`) %>% filter(!is.na(`A. antarctica`)) %>% separate_rows(Orthogroup, `A. antarctica`, convert=T, sep = ', ')
zmar_groups <- groups %>%  select(Orthogroup, `Z. marina`) %>% filter(!is.na(`Z. marina`)) %>% separate_rows(Orthogroup, `Z. marina`, convert=T, sep = ', ')
zmuel_groups <- groups %>%  select(Orthogroup, `Z. muelleri`) %>% filter(!is.na(`Z. muelleri`)) %>% separate_rows(Orthogroup, `Z. muelleri`, convert=T, sep = ', ')
w_australis_groups <- groups %>%  select(Orthogroup, `W. australis`) %>% filter(!is.na(`W. australis`)) %>% separate_rows(Orthogroup, `W. australis`, convert=T, sep = ', ')
s_polyrhiza_groups <- groups %>%  select(Orthogroup, `S. polyrhiza`) %>% filter(!is.na(`S. polyrhiza`)) %>% separate_rows(Orthogroup, `S. polyrhiza`, convert=T, sep = ', ')
l_gibba_groups <- groups %>%  select(Orthogroup, `L. gibba`) %>% filter(!is.na(`L. gibba`)) %>% separate_rows(Orthogroup, `L. gibba`, convert=T, sep = ', ')
rice_groups<- groups %>%  select(Orthogroup, `O. sativa`) %>% filter(!is.na(`O. sativa`)) %>% separate_rows(Orthogroup, `O. sativa`, convert=T, sep = ', ')
```

```{r}
big_joined <- ara_joined %>% mutate(
  present_in_zmuel = case_when(Orthogroup %in% unique(zmuel_groups$Orthogroup) ~ T,
                               TRUE ~ F),
  present_in_zmar = case_when(Orthogroup %in% unique(zmar_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_amphi = case_when(Orthogroup %in% unique(amphi_groups$Orthogroup) ~ T,
                               TRUE ~ F),
  present_in_posi = case_when(Orthogroup %in% unique(posi_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_waustralis = case_when(Orthogroup %in% unique(w_australis_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_spolyrhiza = case_when(Orthogroup %in% unique(s_polyrhiza_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_lgibba = case_when(Orthogroup %in% unique(l_gibba_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_rice = case_when(Orthogroup %in% unique(rice_groups$Orthogroup) ~ T,
                              TRUE ~ F)
  )
```

Done :) Now we have a big table with all *A. thaliana* gene functions and whether these genes are present in clusters of the four seagrasses and three duckweeds.

Let's also add counts - how often is this gene lost?
```{r}
big_joined <- big_joined %>% mutate(Lost_in_Seagrasses = 4 - (present_in_posi + present_in_zmuel + present_in_zmar + present_in_amphi),
                      Lost_in_duckweeds=3 - (present_in_lgibba + present_in_spolyrhiza + present_in_waustralis),
                      Lost_in_both = (Lost_in_Seagrasses + Lost_in_duckweeds))
```

```{r}
big_joined %>% writexl::write_xlsx('data/arabidopsis_gene_level_comparison.xlsx')
big_joined %>% filter_at(vars(starts_with('present')), any_vars(. == FALSE)) %>% writexl::write_xlsx('data/arabidopsis_gene_level_comparison_only_losts.xlsx')
```

That was about loss - can we also check for gene family extension using the *A. thaliana* genes?

# Gene family extension

First, let's count how many members each of these orthogroups has per species.
```{r}
ara_count <- ara_joined %>% count(Orthogroup)
posi_count <- posi_groups %>% count(Orthogroup)
amphi_count <- amphi_groups %>% count(Orthogroup)
zmar_count <- zmar_groups %>% count(Orthogroup)
zmuel_count <- zmuel_groups %>% count(Orthogroup)
lgibba_count <- l_gibba_groups %>% count(Orthogroup)
s_polyrhi_count <- s_polyrhiza_groups %>% count(Orthogroup)
w_aus_count <- w_australis_groups %>% count(Orthogroup)
rice_count <- rice_groups %>% count(Orthogroup)
```

```{r}
counts <- plyr::join_all(list(ara_count, posi_count, amphi_count, zmar_count, zmuel_count, lgibba_count, s_polyrhi_count, w_aus_count, rice_count), by='Orthogroup', type='left') 
names(counts) <- c('Orthogroup', 'A. thaliana', 'P. australis', 'A. antarctica', 'Z. marina', 'Z. muelleri', 'L. gibba', 'S. polyrhiza', 'W. australis', 'O. sativa')
counts
```

```{r}
joined_counts <- left_join(big_joined, counts, by='Orthogroup') %>% select(-starts_with('present')) %>% 
  select(-starts_with('Lost_i'))
joined_counts %>% writexl::write_xlsx('data/arabidopsis_gene_level_counts.xlsx')
```