### 2022-02-21 analysis/plotRgenes.Rmd
better plot

---
title: "R genes"
output: html_notebook
---
```{r setup}
library(tidyverse)
library(wesanderson)
library(ggtree)
library(ggtreeExtra)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

# Visualising R-gene differences here

```{r}
pal <- wes_palette("Zissou1", 3, type = "continuous")

df <- readxl::read_xlsx('./data/R_genes.xlsx')
df %>% ggplot(aes(fill=Class, x=Genome, y=Count)) + geom_bar(stat='identity') +
    scale_fill_manual(values=pal) +  
 coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic"))

```

```{r}
pal <- wes_palette("Zissou1", 4, type = "continuous")

df2 <- df %>% mutate(Class2= case_when(str_detect(Subclass, pattern = 'TN') ~ 'NLR (TNL)',
                                str_detect(Subclass, pattern = '^R') ~ Subclass,
                                TRUE ~ 'NLR (CNL)'))
df2 %>% ggplot(aes(fill=Class2, x=Genome, y=Count)) + geom_bar(stat='identity') +
    scale_fill_manual(values=pal)  + coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic")) 
```

```{r}
df2 %>% dplyr::filter(Genome %in% c('P. australis', 'Z. marina', 'Z. muelleri', 'O. sativa', 'A. antarctica', 'A. thaliana')) %>% 
  ggplot(aes(fill=Class2, x=factor(Genome, levels=c('Z. marina', 'Z. muelleri', 'P. australis', 'A. antarctica', 'O. sativa', 'A. thaliana' )), y=Count)) + 
  geom_bar(stat='identity') +
    scale_fill_manual(values=pal)  + coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic")) +
  xlab('Genome') +
  labs(fill='Class')
```

Let's link those to the phylogeny we got from timetree.org

```{r}
tree <- ape::read.tree('./data/timetree_species.nwk')
tree$tip.label <- c('O. lucimarinus', 'C. reinhardtii', 'P. patens', 'S. moellendorffii', 'O. sativa', 'B. distachyon', 'Z. mays', 'P. australis', 'Z. marina', 'Z. muelleri', 'A. antarctica', 'S. polyrhiza', 'L. gibba', 'V. vinifera', 'A. thaliana', 'S. parvula', 'P. trichocarpa', 'A. trichopada')

```
```{r}
p2 <- df2 %>% ggplot(aes(fill=Class2, x=Genome, y=Count)) + geom_bar(stat='identity') +
    scale_fill_manual(values=pal)  + coord_flip() +  cowplot::theme_minimal_vgrid() +
    theme(axis.text.y = element_text(face = "italic")) 
```

```{r}
p1 <- ggtree(tree)
p1
```

```{r}
df2$label <- df2$Genome
# get species not in tree
subtree <- ape::drop.tip(tree, tree$tip.label[!tree$tip.label %in% df2$label])
```

```{r}
p1 <- ggtree(subtree)
p1 
df3 <- as.data.frame(df2)
df3$label <- df3$Genome
df3$id <- df3$label

```

```{r}
# code from https://thackl.github.io/ggtree-composite-plots

tree_y <-  function(ggtree, data){
  if(!inherits(ggtree, "ggtree"))
    stop("not a ggtree object")
  left_join(select(data, label), select(ggtree$data, label, y)) %>%
    pull(y)
}
# overwrite the default expand for continuous scales
scale_y_tree <- function(expand=expand_scale(0, 0.6), ...){
    scale_y_continuous(expand=expand, ...)
}

# get the range of the ggtree y-axis data
tree_ylim <- function(ggtree){
  if(!inherits(ggtree, "ggtree"))
    stop("not a ggtree object")
  range(ggtree$data$y)
}

# plot data next to a ggtree aligned by shared labels
ggtreeplot <- function(ggtree, data = NULL, mapping = aes(), flip=FALSE,
     expand_limits=expand_scale(0,.6), ...){

  if(!inherits(ggtree, "ggtree"))
    stop("not a ggtree object")

  # match the tree limits
  limits <- tree_ylim(ggtree)
  limits[1] <- limits[1] + (limits[1] * expand_limits[1]) - expand_limits[2]
  limits[2] <- limits[2] + (limits[2] * expand_limits[3]) + expand_limits[4]

  if(flip){
    mapping <- modifyList(aes_(x=~x), mapping)
    data <- mutate(data, x=tree_y(ggtree, data))
    gg <- ggplot(data=data, mapping = mapping, ...) +
      scale_x_continuous(limits=limits, expand=c(0,0))
  }else{
    mapping <- modifyList(aes_(y=~y), mapping)
    data <- mutate(data, y=tree_y(ggtree, data))
    gg <- ggplot(data=data, mapping = mapping, ...) +
      scale_y_continuous(limits=limits, expand=c(0,0))
  }
  gg
}

# get rid of superfluous axis - this works after coord_flip, so it also works
# for the rotated histogram
no_y_axis <- function () 
  theme(axis.line.y = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

```{r}
p3 <-  ggtree(subtree) + geom_tiplab(align=TRUE, fontface='italic') +
  scale_x_continuous(expand=expand_scale(0.8)) + scale_y_tree()

myhist <- ggtreeplot(p3, df3, aes(y=Count, color=Class2, fill=Class2), flip=TRUE) +
  geom_col(aes(fill=Class2,group=Class2,color=Class2)) + 
  #theme(legend.position="none") +
  coord_flip() + no_y_axis()  + 
  theme(legend.position=c(0.6, 0.87)) + 
  labs(fill='Class', color='Class')

p3 + myhist

```