### 2022-02-22 analysis/Orthofinder.Rmd
Add changes

---
title: "Clustering/Orthofinder analysis"
author: "Philipp Bayer"
date: "15 April 2021"
output: html_document
---

Here I use Orthofinder results to pull out genes unique to seagrasses, unique to duckweeds, and make some nice summary tables

```{r setup}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(RColorBrewer)
library(wesanderson)
library(patchwork)
library(UpSetR)
library(kableExtra)
library(ggVennDiagram)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

```{r}
groups <- read_tsv('./data/Orthogroups.tsv.gz')

names(groups) <- c('Orthogroup', 'A. antarctica', 'A. trichopada', 'B. distachyon', 'C. reinhardtii', 'L. gibba', 'O. sativa', 'P. australis', 'P. patens', 'P. trichocarpa', 'S. moellendorffii', 'S. polyrhiza', 'A. thaliana', 'T. parvula', 'V. vinifera', 'Z. mays', 'Z. muelleri', 'Z. marina', 'O. lucimarinus', 'W. australis')
```

```{r}
# for upsetr, we need to know only which OG-groups are shared between species, the actual genes don't matter
per_spec <- groups %>% pivot_longer(-Orthogroup) %>% 
  filter(!is.na(value)) %>% # species not in an orthogroup are still listed, they just have NA genes for this group
  select(-value) # don't need all gene names, speed things up
```

```{r}
# now I want the data in this format:
# listInput <- list(one = c(1, 2, 3, 5, 7, 8, 11, 12, 13), two = c(1, 2, 4, 5, 
#   10), three = c(1, 5, 6, 7, 8, 9, 10, 12, 13))
x <- per_spec %>% 
  select(name, Orthogroup) %>% # turn the table around
  deframe() # convert to named vector
```

```{r}
mylist <- lapply(split(x, names(x)), unname) # yuck - ugly code to convert the named vector to a list
```

```{r}
x <- upset(fromList(mylist), order.by='freq', nsets = length(groups) - 1)
```

```{r fig.width=10}
x
```

Let's get the species-only cluster numbers

```{r}
species_specific_orthos <- per_spec %>% 
  group_by(Orthogroup) %>% 
  summarise(counts = length(name)) %>%
  filter(counts == 1)
```

```{r}
per_spec %>% 
  filter(Orthogroup %in% species_specific_orthos$Orthogroup) %>% 
  group_by(name) %>% 
  count() %>% 
  arrange(n) %>% 
  kbl() %>%
  kable_styling()
```

How many orthogroups are shared between the four seagrasses?

```{r}
newlist <- mylist[c('A. antarctica', 'Z. marina', 'P. australis', 'Z. muelleri')]

x <- upset(fromList(newlist), order.by='freq', nsets = 4)

```

```{r}
x
```

# Connect Orthofinder results with functional table of Arabidopsis genes

File comes from https://www.arabidopsis.org/download_files/Genes/TAIR10_genome_release/TAIR10_functional_descriptions

```{r eval=FALSE}
download.file('https://www.arabidopsis.org/download_files/Genes/TAIR10_genome_release/TAIR10_functional_descriptions', 'data/TAIR10_functional_descriptions')
```
```{r}
functions <- read_tsv('data/TAIR10_functional_descriptions')
```
I also downloaded the gene symbols:

```{r eval=FALSE}
esearch -db gene -query "Arabidopsis thaliana [ORGN]" | esummary | xtract -pattern DocumentSummary -element Name,OtherAliases |  awk -F "\t|," '{OFS="\t"}{print $2,$1}' > arabidopsis_gene_symbols.txt
```

```{r}
symbols <- read_tsv('./data/arabidopsis_gene_symbols.txt', col_names = c('Gene','Symbol'))
```

OK now I need to flip the Orthofinder table to then join the functional table

```{r}
ara_groups <- groups %>% select(Orthogroup, `A. thaliana`) %>% filter(!is.na(`A. thaliana`)) %>% separate_rows(Orthogroup, `A. thaliana`, convert=T)
ara_groups <- ara_groups %>% separate(`A. thaliana`, c('Gene','Number'), remove=FALSE)
ara_groups <- left_join(ara_groups, symbols)
ara_groups$Number <- NULL
head(ara_groups)
```

```{r}
ara_joined <- left_join(ara_groups, functions, by=c(`A. thaliana`='Model_name'))
```

You know what? we shouldn't directly compare a dicot (A. thaliana) with monocots (seagrasses), should also add another regular monocot (rice)

```{r}
posi_groups <- groups %>% select(Orthogroup, `P. australis`) %>% filter(!is.na(`P. australis`)) %>% separate_rows(Orthogroup, `P. australis`, convert=T, sep = ', ')
amphi_groups <- groups %>% select(Orthogroup, `A. antarctica`) %>% filter(!is.na(`A. antarctica`)) %>% separate_rows(Orthogroup, `A. antarctica`, convert=T, sep = ', ')
zmar_groups <- groups %>%  select(Orthogroup, `Z. marina`) %>% filter(!is.na(`Z. marina`)) %>% separate_rows(Orthogroup, `Z. marina`, convert=T, sep = ', ')
zmuel_groups <- groups %>%  select(Orthogroup, `Z. muelleri`) %>% filter(!is.na(`Z. muelleri`)) %>% separate_rows(Orthogroup, `Z. muelleri`, convert=T, sep = ', ')
w_australis_groups <- groups %>%  select(Orthogroup, `W. australis`) %>% filter(!is.na(`W. australis`)) %>% separate_rows(Orthogroup, `W. australis`, convert=T, sep = ', ')
s_polyrhiza_groups <- groups %>%  select(Orthogroup, `S. polyrhiza`) %>% filter(!is.na(`S. polyrhiza`)) %>% separate_rows(Orthogroup, `S. polyrhiza`, convert=T, sep = ', ')
l_gibba_groups <- groups %>%  select(Orthogroup, `L. gibba`) %>% filter(!is.na(`L. gibba`)) %>% separate_rows(Orthogroup, `L. gibba`, convert=T, sep = ', ')
rice_groups<- groups %>%  select(Orthogroup, `O. sativa`) %>% filter(!is.na(`O. sativa`)) %>% separate_rows(Orthogroup, `O. sativa`, convert=T, sep = ', ')
```

```{r}
big_joined <- ara_joined %>% mutate(
  present_in_zmuel = case_when(Orthogroup %in% unique(zmuel_groups$Orthogroup) ~ T,
                               TRUE ~ F),
  present_in_zmar = case_when(Orthogroup %in% unique(zmar_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_amphi = case_when(Orthogroup %in% unique(amphi_groups$Orthogroup) ~ T,
                               TRUE ~ F),
  present_in_posi = case_when(Orthogroup %in% unique(posi_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_waustralis = case_when(Orthogroup %in% unique(w_australis_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_spolyrhiza = case_when(Orthogroup %in% unique(s_polyrhiza_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_lgibba = case_when(Orthogroup %in% unique(l_gibba_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_rice = case_when(Orthogroup %in% unique(rice_groups$Orthogroup) ~ T,
                              TRUE ~ F)
  )
```

Done :) Now we have a big table with all *A. thaliana* gene functions and whether these genes are present in clusters of the four seagrasses and three duckweeds.

Let's also add counts - how often is this gene lost?
```{r}
big_joined <- big_joined %>% mutate(Lost_in_Seagrasses = 4 - (present_in_posi + present_in_zmuel + present_in_zmar + present_in_amphi),
                      Lost_in_duckweeds=3 - (present_in_lgibba + present_in_spolyrhiza + present_in_waustralis),
                      Lost_in_both = (Lost_in_Seagrasses + Lost_in_duckweeds))
```

```{r}
big_joined %>% writexl::write_xlsx('data/arabidopsis_gene_level_comparison.xlsx')
big_joined %>% filter_at(vars(starts_with('present')), any_vars(. == FALSE)) %>% writexl::write_xlsx('data/arabidopsis_gene_level_comparison_only_losts.xlsx')
```

That was about loss - can we also check for gene family extension using the *A. thaliana* genes?

# Gene family extension

First, let's count how many members each of these orthogroups has per species.
```{r}
ara_count <- ara_joined %>% count(Orthogroup)
posi_count <- posi_groups %>% count(Orthogroup)
amphi_count <- amphi_groups %>% count(Orthogroup)
zmar_count <- zmar_groups %>% count(Orthogroup)
zmuel_count <- zmuel_groups %>% count(Orthogroup)
lgibba_count <- l_gibba_groups %>% count(Orthogroup)
s_polyrhi_count <- s_polyrhiza_groups %>% count(Orthogroup)
w_aus_count <- w_australis_groups %>% count(Orthogroup)
rice_count <- rice_groups %>% count(Orthogroup)
```

```{r}
counts <- plyr::join_all(list(ara_count, posi_count, amphi_count, zmar_count, zmuel_count, lgibba_count, s_polyrhi_count, w_aus_count, rice_count), by='Orthogroup', type='left') 
names(counts) <- c('Orthogroup', 'A. thaliana', 'P. australis', 'A. antarctica', 'Z. marina', 'Z. muelleri', 'L. gibba', 'S. polyrhiza', 'W. australis', 'O. sativa')
head(counts)
```

```{r}
joined_counts <- left_join(big_joined, counts, by='Orthogroup') %>% select(-starts_with('present')) %>% 
  select(-starts_with('Lost_i'))
joined_counts %>% writexl::write_xlsx('data/arabidopsis_gene_level_counts.xlsx')
```

# FINAL version redoing - terrestrial vs aquatics vs seagrasses

Plotting all species separately is messy as I have so many species. I'm joining the species into three groups: aquatics, terrestrials, and seagrasses!

```{r}
per_group_spec <-
  per_spec %>% mutate(group = case_when(
    name %in% c(
      'A. trichopada',
      'B. distachyon',
      'O. sativa',
      'P. patens',
      'P. trichocarpa',
      'S. moellendorffii',
      'A. thaliana',
      'T. parvula',
      'V. vinifera',
      'Z. mays'
    ) ~ 'Terrestrials',
    name %in% c(
      'C. reinhardtii',
      'L. gibba',
      'S. polyrhiza',
      'O. lucimarinus',
      'W. australis'
    ) ~ 'Aquatics',
    name %in% c('P. australis', 'Z. muelleri', 'Z. marina', 'A. antarctica') ~ 'Seagrasses'
  )) 
```

```{r}
# now I want the data in this format:
# listInput <- list(one = c(1, 2, 3, 5, 7, 8, 11, 12, 13), two = c(1, 2, 4, 5, 
#   10), three = c(1, 5, 6, 7, 8, 9, 10, 12, 13))
groupx <- per_group_spec %>% 
  select(group, Orthogroup) %>% # turn the table around
  deframe() # convert to named vector
```

```{r}
mylistgroup <- lapply(split(groupx, names(groupx)), unname) # yuck - ugly code to convert the named vector to a list
mylistgroup <- lapply(mylistgroup, unique)
```

```{r}
xgroup <- upset(fromList(mylistgroup), order.by='freq', nsets = length(groups) - 1)
```

```{r fig.width=10}
xgroup
```

MUCH better. The biggest chunk of orthogroups is not in aquatics or seagrasses with 13562 orthogroups. 3437 orthogroups are unique to aquatics, but not in seagrasses - 1183 orthogroups are unique in seagrasses, but not in aquatics. 404 + 3437 + 2217 orthogroups are not in terrestrials.

OK now we need to do the GOenrichment for those three groups.

Let's also make an Venn plot, we have only three groups:
```{r}
p <- ggVennDiagram(mylistgroup, label_alpha=0) + 
  scale_fill_gradientn(colours=wes_palette("Zissou1", 100, type = "continuous")) + 
  theme(legend.position = "none")
```

```{r fig.width=9}
p
```

Nice!

# Write out the per-group GO-terms

Now we pull out all gene-IDs that are only in terrestrials, only in seagrasses, only in aquatics, and unique for each of these three groups, AND we pull out the IDs only present in the four seagrasses each compared with the others.

```{r}
gene_per_spec <- groups %>% pivot_longer(-Orthogroup) %>% 
  filter(!is.na(value)) %>% 
  mutate(group = case_when(
    name %in% c(
      'A. trichopada',
      'B. distachyon',
      'O. sativa',
      'P. patens',
      'P. trichocarpa',
      'S. moellendorffii',
      'A. thaliana',
      'T. parvula',
      'V. vinifera',
      'Z. mays'
    ) ~ 'Terrestrials',
    name %in% c(
      'C. reinhardtii',
      'L. gibba',
      'S. polyrhiza',
      'O. lucimarinus',
      'W. australis'
    ) ~ 'Aquatics',
    name %in% c('P. australis', 'Z. muelleri', 'Z. marina', 'A. antarctica') ~ 'Seagrasses'
  )) 
```