### 2022-02-25 analysis/GOenrichment.Rmd
Replace GO enrichment

---
title: "GO enrichment"
author: "Philipp Bayer"
date: "01/10/2021"
output: html_document
---

```{r setup, include=FALSE}
library(topGO)
library(tidyverse)
library(rvest)
library(wesanderson)
library(httr)
library(stringi)
library(ggrepel)
library(eulerr)
library(UpSetR)
library(rrvgo)
library(kableExtra)
library(patchwork)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Lost in seagrasses

The GO enrichment does not work well on my laptop so I'm setting this to eval=FALSE and run it on a remote server.
The script which writes the input files is a Python script in the scripts/ folder: findClustersUniqueToAquatic.py. This script parses Orthofinder output to pull out genome-specific groups and genes.

```{r eval=FALSE}
# give properly formatted background in format: GO:0005838	GSBRNA2T00088508001;GSBRNA2T00088313001;GSBRNA2T00035842001 
#annAT <- readMappings('BACKGROUND.txt.gz', sep="\t", IDsep=";")
#save(annAT, file='annAtObject.RData')

load('annAtObject.RData')
allgenes <- unique(unlist(annAT))

compare <- function(genelistfile, outname, allgenes, annAT) {
  # give file with your genes of interest, one gene_id per line

  mygenes <-scan(genelistfile ,what="")
  geneList <- factor(as.integer(allgenes %in% mygenes))
  names(geneList) <- allgenes

  GOdata <-new ("topGOdata", ontology = 'BP', allGenes = geneList, nodeSize = 5, annot=annFUN.GO2genes, GO2genes=annAT)
  # using ClassicCount:
  #test.stat <-new ("classicCount", testStatistic = GOFisherTest, name = "Fisher Test")
  #resultsFisherC <-getSigGroups (GOdata, test.stat)
  # using weight01:
  weight01.fisher <- runTest(GOdata, statistic = "fisher")
  # using ClassicCount:
  # allRes <- GenTable(GOdata, classicFisher= resultsFisherC, topNodes = 30)
  # using weight01:
  allRes <- GenTable(GOdata, classicFisher=weight01.fisher,topNodes=30)#,topNodes=100)
  names(allRes)[length(allRes)] <- "p.value"
  p_values <- score(weight01.fisher)
  adjusted_p <- p.adjust(p_values)
  adjusted_p[adjusted_p < 0.05] %>% enframe() %>% write_csv('data/' + outname)

}

```
# Comparing the three big groups

```{r eval=FALSE}
# all genes shared - this runs out of memory
#compare('Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses_and_Terrestrials.txt', 'output/GO_results_genes_in_Aquatics_and_Seagrasses_and_Terrestrials.csv', allgenes, annAT)

# Genes in aquatics and seagrasses - IMPORTANT This is NOT the union - this is the intersection!
# Union is below
compare('Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses.txt', 'output/GO_results_genes_in_Aquatics_and_Seagrasses_NOT_Terrestrials.csv', allgenes, annAT)

# Genes only in aquatics
compare('Lost_present_gene_lists/Genes_only_in_Aquatics.txt', 'output/GO_results_genes_in_Aquatics_NOT_Seagrasses_NOT_Terrestrials.csv', allgenes, annAT)

# Genes only in seagrasses
compare('Lost_present_gene_lists/Genes_only_in_Seagrasses.txt', 'output/GO_results_genes_in_Seagrasses_NOT_Aquatics_NOT_Terrestrials.csv', allgenes, annAT)

# Genes only in terrestrials
compare('Lost_present_gene_lists/Genes_only_in_Terrestrials.txt', 'output/GO_results_genes_in_Terrestrials_NOT_Aquatics_NOT_Seagrasses.csv', allgenes, annAT)

# Genes union of Seagrass and Aquatics
compare('Lost_present_gene_lists/Genes_union_of_Seagrass_and_Aquatics_union.txt', 'output/GO_results_genes_union_Aquatics_and_Seagrasses_NOT_Terrestrials.csv', allgenes, annAT)

```
# Seagrass only comparisons

Now we compare seagrasses within each other.

For the seagrass-only comparisons, I'm using a Seagrass-only background as that makes more biological sense to me.s

```{r eval=FALSE}
# give properly formatted background in format: GO:0005838	GSBRNA2T00088508001;GSBRNA2T00088313001;GSBRNA2T00035842001 
#sannAT <- readMappings('SEAGRASSBACKGROUND.txt', sep="\t", IDsep=";")
#save(sannAT, file='sannAtObject.RData')

load('sannAtObject.RData')
sallgenes <- unique(unlist(sannAT))

```
# lost in seagrasses

```{r eval=FALSE}
compare('Lost_present_gene_lists/Genes_lost_in_A_antarctica_not_other_seagrasses.txt', 'output/GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv', sallgenes, sannAT)

compare('Lost_present_gene_lists/Genes_lost_in_Z_marina_not_other_seagrasses.txt', 'output/GO_results_genes_lost_Z_marina_not_other_seagrasses.csv', sallgenes, sannAT)

compare('Lost_present_gene_lists/Genes_lost_in_P_australis_not_other_seagrasses.txt', 'output/GO_results_genes_lost_P_australis_not_other_seagrasses.csv', sallgenes, sannAT)

compare('Lost_present_gene_lists/Genes_lost_in_Z_muelleri_not_other_seagrasses.txt', 'output/GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv', sallgenes, sannAT)
```

# Present in seagrasses

```{r eval=FALSE}
compare('Lost_present_gene_lists/Genes_only_in_Z_marina_not_other_seagrasses.txt', 'output/GO_results_genes_only_Z_marina_not_other_seagrasses.csv', sallgenes, sannAT)

compare('Lost_present_gene_lists/Genes_only_in_Z_muelleri_not_other_seagrasses.txt', 'output/GO_results_genes_only_Z_muelleri_not_other_seagrasses.csv', sallgenes, sannAT)

compare('Lost_present_gene_lists/Genes_only_in_P_australis_not_other_seagrasses.txt', 'output/GO_results_genes_only_P_australis_not_other_seagrasses.csv', sallgenes, sannAT)

compare('Lost_present_gene_lists/Genes_only_in_A_antarctica_not_other_seagrasses.txt', 'output/GO_results_genes_only_A_antarctica_not_other_seagrasses.csv', sallgenes, sannAT)
```

Alright now we have all these different GO terms in all these files - we can send them to revigo for visualiation and some deduplication!

# Revigo

This code is based on http://revigo.irb.hr/CodeExamples/revigo.R.txt

```{r}
results_list <- list()

for (f in list.files('./output/', pattern='*.csv')){
  filename <- paste('./output/', f, sep='')
  go_and_pvalues <- readChar(filename, file.info(filename)$size)
  go_and_pvalues <- gsub(',', ' ', go_and_pvalues)

  httr::POST(
    url = "http://revigo.irb.hr/StartJob.aspx",
    body = list(
      cutoff = "0.7",
      valueType = "pvalue",
     # speciesTaxon = "4577", # zea mays
     #speciesTaxon = '39947', # japonica
     speciesTaxon = '3702', # arabidopsis
      measure = "SIMREL",
      goList = go_and_pvalues
    ),
    # application/x-www-form-urlencoded
    encode = "form"
  ) -> res

  dat <- httr::content(res, encoding = "UTF-8")

  jobid <- jsonlite::fromJSON(dat,bigint_as_char=TRUE)$jobid

  # Check job status
  running <- "1"
  while (running != "0" ) {
      httr::POST(
        url = "http://revigo.irb.hr/QueryJobStatus.aspx",
        query = list( jobid = jobid )
      ) -> res2
      dat2 <- httr::content(res2, encoding = "UTF-8")
      running <- jsonlite::fromJSON(dat2)$running
      Sys.sleep(1)
  }

  # Fetch results
  httr::POST(
    url = "http://revigo.irb.hr/ExportJob.aspx",
    query = list(
      jobid = jobid, 
      namespace = "1",
      type = "CSVTable"
    )
  ) -> res3

  dat3 <- httr::content(res3, encoding = "UTF-8")

  dat3 <- stri_replace_all_fixed(dat3, "\r", "")
  # Now we have a csv table in a string!

  # read_csv does not like the ', ', it wants ','
  dat <- read_csv(gsub(', ', ',', dat3), show_col_types = FALSE)

  # do we even have results?
  if(nrow(dat) == 0){next}
  results_list[[f]] <- dat

}
```

OK we have a list with all results in a big list. Now we can plot!

```{r warning=FALSE, echo=FALSE, message=FALSE}
# lots of warnings and messages here so I'm hiding these
plot_list <- list()
for (f in names(results_list)) {
  dat <- results_list[[f]]
  if(nrow(dat) == 0) {next}

  names(dat) <- c("term_ID","description","frequency", 'plot_X', 'plot_Y', 'log_size', 'value', 'uniqueness', 'dispensability', 'eliminated', 'representative')
  one.data <- dat
  if (one.data[1,] == '<html>') {next} # some datasets result in error
  one.data <- one.data [(one.data$plot_X != "null" & one.data$plot_Y != "null") & (! is.na(one.data$frequency)) & (one.data$value != 'null'), ];
  one.data <- suppressMessages(type_convert(one.data)) # guess data types, but also: shush
  top_ten_values <- one.data %>% arrange(value) %>% head(10) %>% pull(term_ID) 
  one.data <- one.data %>% mutate(description2 = case_when(
    term_ID %in% top_ten_values ~ description, 
    TRUE ~ ''))
  p1 <-
    ggplot(data = one.data, aes(plot_X, plot_Y, label = description2)) +
    geom_point(aes(colour = value, size = log_size), alpha = I(0.6)) +
    scale_size_area() +
    scale_colour_gradientn(colours = rev(wes_palette("Zissou1", 100, type = "continuous")), 
                           limits = c(min(one.data$value), 0)) +
    geom_point(
      aes(plot_X, plot_Y, size = log_size),
      shape = 21,
      fill = "transparent",
      colour = I (alpha ("black", 0.6))
    ) +
    scale_size_area() +
    geom_label_repel(max.overlaps = Inf, box.padding = 0.5,
                     aes(point.size = log_size), alpha=0.9) +
    theme_minimal() +
    labs(color = 'log(p-value)',
         size = 'Frequency') +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    ) +
    NULL
    #ggtitle(f)

  plot_list[[f]] <- p1

}
```

# Making Supplementary Tables for the REVIGO results

Let's also pull these terms out as tables. `Value` is the unadjusted p-value (-3.7 = 10 * -3.7)

```{r results='asis'}
for( i in names(results_list)) {
  print(i)
  print(results_list[[i]] %>% filter(Eliminated == FALSE) %>% select(TermID, Name, Value) %>% arrange(Value) %>% kbl() %>%  kable_styling())
  cat("\n")
}

```

Let's have a look at all of these plots. Manually zooming in leads to ggrepel reloading labels, so on the small scale a lot of these plots don't have labels.
```{r}
for(i in names(plot_list)) {
  file_name = paste('output/',  i, '.png', sep='')

  cowplot::save_plot(file_name, plot_list[[i]], base_height=6)

}
```

Let me also try and make some patchworks.

```{r fig.width=10}
# For some reason with this plot the legends aren't correctly collected by patchwork, 
# so I need to manually turn one off
patchy <- (plot_list[["GO_results_genes_in_Terrestrials_NOT_Aquatics_NOT_Seagrasses.csv"]] + theme(legend.position = "none")) /
  plot_list[["GO_results_genes_union_Aquatics_and_Seagrasses_NOT_Terrestrials.csv"]] + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A')
patchy
cowplot::save_plot('output/GO_results_terrestrials_vs_union_aquatics_seagrasses.png', patchy, base_height=7)
```

OK I'll better make this figure for the seagrasses alone, it's quite messy this way.

# Seagrass-specific loss

Let's check which GO terms overlap between the 4 seagrasses!

How many shared lost GO-terms are there?

```{r}
a <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$Name,
          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$Name,
          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$Name,
          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$Name)

a_go_ids <- list(`P. australis` = results_list$GO_results_genes_lost_P_australis_not_other_seagrasses.csv$TermID,
          `A. antarctica` = results_list$GO_results_genes_lost_A_antarctica_not_other_seagrasses.csv$TermID,
          `Z. marina` = results_list$GO_results_genes_lost_Z_marina_not_other_seagrasses.csv$TermID,
          `Z. muelleri` = results_list$GO_results_genes_lost_Z_muelleri_not_other_seagrasses.csv$TermID)
```

```{r}
plot(venn(a), 
     quantities = TRUE,
     fill = rev(wes_palette("Zissou1", 15, type = 'continuous')),
    alpha = 0.5,
     labels = list(font = 4))
```

```{r}
upset(fromList(a), order.by='freq', )
```

```{r}
a_no_ara <- list(`P. australis` = results_list$missing_posi_vs_all_GO.txt$Name,
          `A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name)
```

What are the shared GO-terms in seagrasses, WITHOUT the Ara loss??
```{r}
Reduce(union, a) %>% enframe() %>% writexl::write_xlsx('./output/Seagrasses_shared_lost_genes.xlsx')
```

What if we remove Posidonia?

```{r}
b <- list(`A. antarctica` = results_list$missing_amphi_vs_all_GO.txt$Name,
          `Z. marina` = results_list$missing_zmar_vs_all_GO.txt$Name,
          `Z. muelleri` = results_list$missing_zmuel_vs_all_GO.txt$Name)
intersections <- Reduce(intersect, b)
intersections[grepl('ethylene', intersections)]
```
OK we need a big list of all GO-terms here - which GO-term is lost in which species. That will be a supplementary table.

```{r}
all_species <- c("P. australis","A. antarctica","Z. marina","Z. muelleri", 'A. thaliana')  
all_go_terms <- Reduce(union, a)
all_go_ids <- Reduce(union, a_go_ids)
results_d <- data.frame('GOID' = character(),
                        'GO' = character(),
                        'P. australis' = character(),
                        'A. antarctica' = character(),
                        'Z. marina' = character(),
                        'Z. muelleri' = character(),
                        'A. thaliana' = character())

for (index in seq_along(all_go_terms)){
  go <- all_go_terms[index]
  go_id <- all_go_ids[index]
  specs <- c()
  for (species in names(a)) {
    if ( length(a[[species]][grep(paste('^', go, '$', sep=''), a[[species]])]) > 0 ) {
      specs <- c(specs, species)
    }
  }

  results_d[index,] <- c(go_id, go, gsub('FALSE', 'Present', gsub('TRUE', 'Lost', all_species %in% specs)))
}
```

```{r}
writexl::write_xlsx(results_d, 'output/Lost_GO_terms_in_five_species.xlsx')
```

# After filtering for plant-specific GO-terms

We will use the GO-terms that are plant-specific as identified by the GOMAP paper.
See https://github.com/wkpalan/GOMAP-maize-analysis/blob/main/6.plantSpecific/1.getSppSpecific.R
or https://plantmethods.biomedcentral.com/articles/10.1186/s13007-021-00754-1

```{r}
go_plant <- read_tsv('https://raw.githubusercontent.com/wkpalan/GOMAP-maize-analysis/main/data/go/speciesSpecificGOTerms.txt')
```

```{r}
# taxon 33090 is Viridiplantae
plantSpecificGO <- go_plant %>% dplyr::filter(`NCBITaxon:33090`==1) %>% pull(GOterm)
plantSpecificGO <- c(plantSpecificGO,c("GO:0005575","GO:0008150","GO:0003674"))
```

```{r}
results_d %>% filter(GOID %in% plantSpecificGO) %>% writexl::write_xlsx('output/Lost_GO_terms_in_five_species.PlantSpecific.xlsx')
```

Now let's redo the Venn diagram with those filtered GO-terms

```{r}
filters <- lapply(a_go_ids, function(ch) ch %in% plantSpecificGO)

newa <- list()
for (species in names(filters)) {
  before <- a[[species]]
  after <- before[filters[[species]]]
  newa[[species]] <- after
}

```

```{r}
plot(venn(newa), 
     quantities = TRUE,
     fill = rev(wes_palette("Zissou1", 15, type = 'continuous')),
    alpha = 0.5,
     labels = list(font = 4))
```

Not much difference?