### 2022-03-09 analysis/index.Rmd
add better groups

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

This website collects comparisons of *Amphibolis* and *Posidonia* comparative analyses.

[Repeat comparison plots](plot.html) - Comparing genome sizes and repeat/TE content

[Phylogenies](OTT.html) - plotting just phylogenies

[Clustering results](Orthofinder.html) - analysing Orthofinder clustering results

[R gene plotting](plotRgenes.html) - making phylogenies and showing R-gene number differences

[Gene Loss comparison by reads](GeneLossComparisonReads.html) - making heatmaps and barcharts to compare read alignments

[GO enrichment](GOenrichment.html) - running topGO and REVIGO to make nice plots of GO-enrichment

[GO and Ortho](GO_and_ortho.html) - merging plots from GO enrichment and Orthofinder analysis