### 2022-03-10 analysis/GO_and_ortho.Rmd
GO-term updates! more groups! for the price of one!

---
title: "Visualising gene differences based on read alignments"
author: "Philipp Bayer"
date: "9th March 2022"
output: html_document
---

```{r setup}
library(tidyverse)
library(patchwork)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

In this one I just merge some plots from the Orthofinder and GO enrichment Rmds, which I have saved as Rdata objects.

```{r}
load('output/all_GO_plots.Rdata')
load('output/group_venn_image.Rdata')
load('output/seagrass_venn_image.Rdata')
```
```{r}
group_venn
```
```{r}
seagrass_venn
```

OK now we can make patchworks with these Venn diagrams and their associated GO-terms
```{r fig.width=8, fig.height=10}
sum1 <- group_venn / (plot_list$GO_results_genes_in_Terrestrials_NOT_Aquatics_NOT_Seagrasses.csv + theme(legend.position="bottom")) +  plot_annotation(tag_levels = 'A') +plot_layout(heights=c(1.5,1))
sum1
cowplot::save_plot(sum1, filename = 'output/patchwork_terrestrials_gene_loss.png', base_height = 10, base_width = 8)
```

```{r fig.width=8, fig.height=10}
sum2 <- seagrass_venn / (plot_list$GO_results_genes_in_all_seagrasses_vs_seagrassesBackground_intersect.csv + theme(legend.position="bottom")) +  plot_annotation(tag_levels = 'A') +plot_layout(heights=c(1.5,1))
sum2
cowplot::save_plot(sum2, filename = 'output/patchwork_seagrass_gene_loss.png', base_height = 10, base_width = 8)
```