### 2022-03-11 analysis/GeneLossComparisonReads.Rmd
finalised MSA!

---
title: "Visualising gene differences based on read alignments"
author: "Philipp Bayer"
date: "21 February 2022"
output: html_document
---

Here I compare read alignment differences with known Arabidopsis genes.

```{r setup, message=FALSE}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(RColorBrewer)
library(wesanderson)
library(gghighlight)
library(patchwork)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

The data is based on Supplementary Table S6 and S7.

# Stomata genes

```{r}
stomata <- tibble::tribble(
             ~Gene.symbol, ~A..antarctica, ~P..australis, ~H..ovalis, ~Z..muelleri, ~Z..marina,
                 "SBT1.2",              0,             0,                  5.63,          8.2,       4.34,
                  "EPFL9",              0,             0,                     0,            0,          0,
                   "EPF1",              0,         25.08,                     0,            0,          0,
                    "TMM",           3.35,             0,                     0,            0,          0,
                   "EPF2",              0,             0,                     0,            0,          0,
                  "MYB88",              0,             0,                  6.74,         2.34,          0,
                   "MUTE",              0,             0,                     0,            0,          0,
                   "SPCH",           7.95,          3.29,                  5.84,         4.47,          0,
                   "FAMA",           6.99,         20.64,                  6.83,         4.98,       4.98,
                  "SCRM2",           21.8,         25.57,                 19.66,        26.16,      22.25,
                    "FLP",            3.2,          8.77,                  4.42,            0,          0
             )

```

```{r}
stomata_long <- stomata %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
  filter(Species != 'H. ovalis')
```

```{r}
stom_p <- stomata_long %>% ggplot(aes(x=forcats::fct_rev(Species), y=Coverage, fill=Species, color=Species)) + 
  geom_boxplot(
    aes(fill = after_scale(colorspace::lighten(fill, .9))),
    size = 1.5, outlier.shape = NA
  ) +
  geom_jitter(width = .1, size = 4, alpha = .5) +
  ylim(c(-1,100)) +

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage (%)') +
  coord_flip() + xlab('Species')

stom_p
```

```{r}
label <- stomata_long #%>% mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) 
label <- label %>% 
  mutate(mylabel = case_when(Coverage >= 100 ~ '',
                             Coverage == 0 ~ '',
                                    TRUE ~ as.character(Coverage)))

stom_heat <- stomata_long %>% 
  #mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
  geom_tile(alpha=0.9, color='white') + 
  #geom_text(aes(label=mylabel), data=label, size=3.5) + 
  theme(axis.text.y = element_text(face = "italic"),
        axis.text.x = element_text(face='italic'),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        #legend.position = "none"
        ) +
  #scale_fill_continuous(limits=c(0,100)) +
  scale_fill_distiller(palette='Spectral', limits=c(0,100))#palette= 'Rdbu', limits=c(0,100), type = 'discrete')

stom_heat
```

# Ethylene

```{r}
ethylene <- tibble::tribble(
              ~Gene.symbol, ~A..antarctica, ~P..australis, ~H..ovalis, ~Z..muelleri, ~Z..marina,
                    "ACO1",              0,         53.16,          0,            0,          0,
                    "ACO2",              0,         11.53,          0,            0,          0,
                    "ACO4",              0,         15.74,          0,            0,          0,
                    "ACO5",              0,             0,          0,         6.39,        5.3,
                    "ACS1",           9.34,          5.11,          0,            0,          0,
                    "ACS2",           7.24,          5.97,          0,         4.43,          0,
                    "ACS4",          18.74,         25.19,          0,         8.49,       8.21,
                    "ACS5",          19.25,         50.39,       4.46,            0,          0,
                    "ACS6",              0,             0,          0,            0,          0,
                    "ACS7",          10.27,         66.29,          0,         6.47,          0,
                    "ACS8",          38.58,         42.62,       6.24,            0,          0,
                    "ACS9",          20.52,         36.94,          0,            0,          0,
                   "ACS11",              0,         22.05,       7.45,            0,          0,
                    "ERS1",          35.23,         40.88,          0,          2.5,          0,
               "ETR1/EIN1",          49.57,         51.87,          0,         3.83,          0,
                    "ETR2",              0,         21.96,          0,            0,          0,
                    "EIN4",           1.74,         31.33,          0,            0,       1.78,
                    "CTR1",          26.68,         31.55,       5.96,            0,          0,
                    "EIN2",              0,         14.54,      11.12,            0,          0,
                    "EBF1",              0,             0,        4.5,            0,          0,
                    "EBF2",              0,         12.45,       4.22,            0,          0,
                    "EIN3",          27.03,         43.08,      36.83,        31.27,      31.11,
                    "MYC2",           4.43,         27.83,      10.31,        23.61,       23.5,
                    "MYC3",             19,         17.14,      10.06,        12.14,      13.77,
                    "MYC4",              0,          3.84,       8.19,        20.96,      11.75
              )

```

```{r}
ethylene_long <- ethylene %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
  filter(Species != 'H. ovalis')
```

```{r}
eth_p <- ethylene_long %>% ggplot(aes(x=forcats::fct_rev(Species), y=Coverage, fill=Species, color=Species)) + 
  geom_boxplot(
    aes(fill = after_scale(colorspace::lighten(fill, .9))),
    size = 1.5, outlier.shape = NA
  ) +
  geom_jitter(width = .1, size = 4, alpha = .5) +
  ylim(c(-1,100)) +

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage (%)') +
  coord_flip() +
  xlab('Species')
eth_p
```

```{r}
eth_heat <- ethylene_long %>% 
  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
  #mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
  geom_tile() + 
  ylab('Gene') +
  theme(axis.text.y = element_text(face = "italic"),
        axis.text.x = element_text(face='italic'),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
       #legend.position = "none"
       ) +
  #scale_fill_continuous(limits=c(0,100)) +
  scale_fill_distiller(palette='Spectral', limits=c(0,100))#palette= 'Rdbu', limits=c(0,100), type = 'discrete')

eth_heat
```

# NDH genes

```{r}
ndh <- tibble::tribble(
         ~Gene.symbol, ~A..antarctica, ~P..australis, ~H..ovalis, ~Z..muelleri, ~Z..marina,
               "ndhL",              0,             0,          0,        16.67,      16.67,
               "ndhM",              0,             0,          0,        36.24,       34.4,
               "ndhN",          48.73,             0,          0,        66.35,      59.68,
               "ndhO",              0,             0,          0,        33.54,      22.22,
               "ndhS",          12.35,         28.29,          0,         34.4,      30.15,
               "ndhT",              0,          8.67,          0,        19.47,      16.67,
               "ndhU",              0,             0,          0,         9.89,          0,
              "PnsB1",           4.26,         42.42,          0,         32.4,      31.24,
              "PnsB2",              0,             0,          0,            0,          0,
              "PnsB3",              0,          8.46,          0,        16.26,       9.76,
              "PnsB4",              0,         19.13,          0,        18.75,      18.75,
              "PnsL1",              0,             0,          0,            0,          0,
              "PnsL2",          16.58,             0,          0,        45.03,      15.01,
              "PnsL3",              0,             0,          0,            0,          0,
              "PnsL4",           7.19,         23.55,          0,        31.04,      45.11,
              "PnsL5",          52.82,         62.82,      57.31,        48.72,      46.28,
               "CRR6",              0,             0,          0,        43.86,      33.06,
               "CRR7",              0,             0,          0,            0,          0,
              "Lhca5",              0,         35.15,          0,        18.29,      27.76,
              "Lhca6",              0,         47.85,          0,        58.43,      60.27,
              "CRR27",           21.9,           9.2,       9.26,        36.38,      41.78,
              "CRR41",              0,             0,          0,        44.03,       7.86,
               "PGR5",          55.97,         56.72,      53.98,        55.22,      55.22,
             "PGRL1A",          17.54,         23.79,      29.54,        33.54,      31.08,
               "CRR2",          41.49,          61.4,       2.94,        23.51,      13.93,
               "CRR3",              0,             0,          0,            0,          0,
              "CRR42",              0,             0,          0,        30.89,      28.44,
               "PQL3",              0,             0,          0,            0,          0,
               "NDF5",              0,             0,          0,            0,          0
         )

```

```{r}
ndh_long <- ndh %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
  filter(Species != 'H. ovalis')
```

```{r}
ndh_p <- ndh_long %>% ggplot(aes(x=forcats::fct_rev(Species), y=Coverage, fill=Species, color=Species)) + 
  geom_boxplot(
    aes( fill = after_scale(colorspace::lighten(fill, .9))),
    size = 1.5, outlier.shape = NA
  ) +
  geom_jitter(width = .1, size = 4, alpha = .5) +
  ylim(c(-1,100)) +

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Coverage %') +
  coord_flip() +
  xlab('Species')
ndh_p
```
```{r fig.width=10}
label <- ndh_long
label <- label %>% 
  mutate(mylabel = case_when(Coverage >= 100 ~ '',
                             Coverage == 0 ~ '',
                                    TRUE ~ as.character(Coverage)))

ndh_heat <- ndh_long %>% 
  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
  #mutate(Gene.symbol = stringr::str_to_lower(Gene.symbol)) %>% 
  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
  geom_tile() + 
  geom_tile(alpha=0.9, color='white') + 
  #geom_text(data=label, aes(label=mylabel), size=3.5) +
  theme(axis.text.y = element_text(face = "italic"),
        axis.text.x = element_text(face = 'italic'),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        #legend.position = "none"
        )  +
  #scale_fill_continuous(limits=c(0,100)) +
  scale_fill_distiller(palette='Spectral', limits=c(0,100))#palette= 'Rdbu', limits=c(0,100), type = 'discrete')

ndh_heat
```

The following data is slightly different because it compares lengths of annotated genes in the chloroplast assembly.

```{r}
ndh_chloro <- tibble::tribble(
  ~Gene.symbol, ~A..antarctica, ~H..ovalis, ~P..australis, ~Z..marina, ~Z..muelleri,
        "ndhA",              0,          0,             0,      101.8,        101.4,
        "ndhB",              0,          0,             0,      100.4,        100.9,
      "ndhB-2",              0,          0,             0,      100.4,        100.9,
        "ndhC",           92.5,          0,           100,        100,          100,
        "ndhD",           29.4,          0,             0,        100,          100,
        "ndhE",              0,          0,          45.4,        100,          100,
        "ndhF",              0,          0,             0,       98.4,         98.3,
        "ndhG",           45.9,          0,             0,        100,          100,
        "ndhH",             34,       63.1,          63.8,        100,          100,
        "ndhI",              0,          0,             0,      110.5,        110.5,
        "ndhJ",           70.8,          0,             0,        100,          100,
        "ndhK",             55,          0,         109.3,      109.8,        109.8
  )

```

```{r}
ndh_chl_long <- ndh_chloro %>% pivot_longer(-Gene.symbol, values_to = 'Coverage', names_to='Species') %>% 
  mutate(Species = gsub('\\.\\.', '. ', Species)) %>% 
  filter(Species != 'H. ovalis')
```

Stylistic choice - if the new gene is longer than the Arabidopsis one, set to 100%.

```{r}
ndh_chl_p <- ndh_chl_long %>%
  mutate(Coverage = case_when(Coverage > 100 ~ 100,
                              TRUE ~ Coverage)) %>% 
  ggplot(aes(x=forcats::fct_rev(Species), y=Coverage, fill=Species, color=Species)) + 
  geom_boxplot(
    aes( fill = after_scale(colorspace::lighten(fill, .9))),
    size = 1.5, outlier.shape = NA
  ) +
  geom_jitter(width = .1, size = 4, alpha = .5) +
  ylim(c(-1,100.1)) +

  theme(legend.position = "None",
  axis.text.y = element_text(face = "italic")) +
  scale_color_brewer(palette='Dark2') +
  ylab('Gene size similarity (%)') +
  xlab('Species') +
  coord_flip()
ndh_chl_p
```

```{r}
label <- ndh_chl_long %>% filter(Gene.symbol != 'ndhB-2')
label <- label %>% 
  mutate(mylabel = case_when(Coverage >= 100 ~ '',
                             Coverage == 0 ~ '',
                                    TRUE ~ as.character(Coverage)))

ndh_chl_heat <- ndh_chl_long %>% 
  dplyr::filter(Gene.symbol != 'ndhB-2') %>% 
  mutate(Coverage = case_when(Coverage > 100 ~ 100,
                              TRUE ~ Coverage)) %>% 
  ggplot(aes(x=Species, y=factor(Gene.symbol, level=rev(unique(Gene.symbol))), fill=Coverage)) + 
  geom_tile(alpha=0.9, color='white') + 
  #geom_text(data=label, aes(label=mylabel), size=3.5) +
  ylab('Gene') +
  theme(axis.text.y = element_text(face = "italic"),
        axis.text.x = element_text(face = 'italic'),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        #legend.position = "none"
        ) +
#  scale_fill_continuous(limits=c(0,100)) +
    scale_fill_distiller(palette='Spectral', limits=c(0,100))#palette= 'Rdbu', limits=c(0,100), type = 'discrete')

ndh_chl_heat
```

# all together

```{r fig.width=10}
patchwork <- (eth_p + stom_p) / ( ndh_chl_p + ndh_p )
# patchwork[[2]] = patchwork[[2]] + theme(axis.text.y = element_blank(),
#                                         axis.ticks.y = element_blank(),
#                                         axis.title.y = element_blank() )
# patchwork[[1]] = patchwork[[1]] + theme(axis.text.y = element_blank(),
#                                         axis.ticks.y = element_blank(),
#                                         axis.title.y = element_blank() )
patchwork+ plot_annotation(tag_levels = "A")
```

```{r fig.height=7}
patchwork <- (eth_p + ndh_chl_heat)/
 (ndh_p + stom_heat  )
#patchwork[[2]] = patchwork[[2]] + theme(axis.title.x = element_blank() )
#patchwork[[1]] = patchwork[[1]] + theme(axis.title.x = element_blank() )
puh <- patchwork+ plot_annotation(tag_levels = "A") + plot_layout(guides='collect')
puh
cowplot::save_plot(filename = 'output/patched_gene_loss.png', plot = puh, base_height = 7)
```

Hmmmmmmmmm

```{r fig.width=10}

patchwork <- (eth_heat + ndh_chl_heat)/
 ( ndh_heat + stom_heat  ) & theme(axis.text = element_text(size = 8))
#patchwork[[2]] = patchwork[[2]] + theme(axis.title.x = element_blank() )
#patchwork[[1]] = patchwork[[1]] + theme(axis.title.x = element_blank() )
puh <- patchwork+ plot_annotation(tag_levels = "A") + plot_layout(guides='collect')
puh
cowplot::save_plot(filename = 'output/patched_gene_loss_heatmaps.png', plot = puh, base_height = 8)
```