### 2022-03-15 analysis/ThreeDStructure.Rmd
wflow_publish("analysis/ThreeDStructure.Rmd")

---
title: "Visualising 3D structures"
author: "Philipp Bayer"
date: "15 March 2022"
output: html_document
---

Let's compare some 3D structures from the interesting genes in [MSA](MSA.html)

```{r message=FALSE, warning=FALSE}
library(bio3d)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

# SOS3

First, let's just load one of them 
```{r}
# the 'real' one from X-ray crystallography
ara_SOS3_real <- read.pdb('data/SOS3_OG0000189/Arabidopsis_SOS3_1v1f.pdb')
# the alphafold ones

amphibolis_SOS3_alpha <- read.pdb('data/SOS3_OG0000189/Amphibolis_fixed_SOS3_40588.result/Amphibolis_fixed_SOS3_40588_unrelaxed_rank_1_model_4.pdb')
posi_SOS3_alpha <- read.pdb('data/SOS3_OG0000189/6P_australis_SOS3_dcfd4.result/6P_australis_SOS3_dcfd4_unrelaxed_rank_1_model_1.pdb')
```

```{r}
ara_SOS3_gaps <- gap.inspect(ara_SOS3_real$xyz)
ara.xyz <- ara_SOS3_real$xyz[ara_SOS3_gaps$f.inds]

amphi_gaps <- gap.inspect(amphibolis_SOS3_alpha$xyz)
amph.xyz <- amphibolis_SOS3_alpha$xyz[amphi_gaps$f.inds]

a <- torsion.xyz(ara.xyz, atm.inc=1)
b <- torsion.xyz(amph.xyz, atm.inc=1)
d.ab <- wrap.tor(a-b)
d.ab[is.na(d.ab)] <- 0
plot.bio3d(abs(d.ab), typ="h", xlab="Residue No.", 
           ylab = "Difference Angle")

```
```{r}
posi_SOS3_gaphs <- gap.inspect(posi_SOS3_alpha$xyz)
posi.xyz <- posi_SOS3_alpha$xyz[posi_SOS3_gaphs$f.inds]

a <- torsion.xyz(posi.xyz, atm.inc=1)
b <- torsion.xyz(amph.xyz, atm.inc=1)
d.ab <- wrap.tor(a-b)
d.ab[is.na(d.ab)] <- 0
plot.bio3d(abs(d.ab), typ="h", xlab="Residue No.", 
           ylab = "Difference Angle")

```

```{r}
library(r3dmol)

m_glimpse(ara_SOS3_real, outline = FALSE) %>% 
  m_spin()
```

```{r}

m_glimpse(posi_SOS3_alpha, outline = FALSE) %>% 
  m_spin()
```

```{r}
m1 <- r3dmol() %>%
  m_add_model(data = posi_SOS3_alpha, format = "pdb") %>%
  m_zoom_to()
m2 <- r3dmol() %>%
  m_add_model(data = amphibolis_SOS3_alpha, format = "pdb") %>%
  m_zoom_to()
m3 <- r3dmol() %>%
  m_add_model(data = '1v1f', format = "pdb") %>%
  m_zoom_to()
m_grid(
  viewer = list(m1, m2, m3),
  rows = 2,
  cols = 2,
  #control_all = TRUE,
  #viewer_config = m_viewer_spec(
  #  backgroundColor = "lightblue"
  #)
)
```