### 2022-04-26 analysis/Orthofinder.Rmd
start to remove gibba

---
title: "Clustering/Orthofinder analysis"
author: "Philipp Bayer"
date: "15 April 2021"
output: html_document
---

Here I use Orthofinder results to pull out genes unique to seagrasses, unique to duckweeds, and make some nice summary tables

```{r setup, message=FALSE}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot())
library(RColorBrewer)
library(wesanderson)
library(patchwork)
library(UpSetR)
library(kableExtra)
library(ggVennDiagram)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

```{r}
groups <- read_tsv('./data/Orthogroups.tsv.gz')

names(groups) <- c('Orthogroup', 'A. antarctica', 'A. trichopada', 'B. distachyon', 'C. reinhardtii', 'L. gibba', 'O. sativa', 'P. australis', 'P. patens', 'P. trichocarpa', 'S. moellendorffii', 'S. polyrhiza', 'A. thaliana', 'T. parvula', 'V. vinifera', 'Z. mays', 'Z. muelleri', 'Z. marina', 'O. lucimarinus', 'W. australis')
```

Remove L. gibba

```{r}
groups$`L. gibba` <- NULL
```

```{r}
# for upsetr, we need to know only which OG-groups are shared between species, the actual genes don't matter
per_spec <- groups %>% pivot_longer(-Orthogroup) %>% 
  filter(!is.na(value)) %>% # species not in an orthogroup are still listed, they just have NA genes for this group
  select(-value) # don't need all gene names, speed things up
```

```{r}
# now I want the data in this format:
# listInput <- list(one = c(1, 2, 3, 5, 7, 8, 11, 12, 13), two = c(1, 2, 4, 5, 
#   10), three = c(1, 5, 6, 7, 8, 9, 10, 12, 13))
x <- per_spec %>% 
  select(name, Orthogroup) %>% # turn the table around
  deframe() # convert to named vector
```

```{r}
mylist <- lapply(split(x, names(x)), unname) # yuck - ugly code to convert the named vector to a list
```

```{r}
x <- upset(fromList(mylist), order.by='freq', nsets = length(groups) - 1)
```

```{r fig.width=10}
x
```

Let's get the species-only cluster numbers

```{r}
species_specific_orthos <- per_spec %>% 
  group_by(Orthogroup) %>% 
  summarise(counts = length(name)) %>%
  filter(counts == 1)
```

```{r}
per_spec %>% 
  filter(Orthogroup %in% species_specific_orthos$Orthogroup) %>% 
  group_by(name) %>% 
  count() %>% 
  arrange(n) %>% 
  kbl() %>%
  kable_styling()
```

How many orthogroups are shared between the four seagrasses?

```{r}
newlist <- mylist[c('A. antarctica', 'Z. marina', 'P. australis', 'Z. muelleri')]

x <- upset(fromList(newlist), order.by='freq', nsets = 4)

```

```{r}
x
```

# Connect Orthofinder results with functional table of Arabidopsis genes

File comes from https://www.arabidopsis.org/download_files/Genes/TAIR10_genome_release/TAIR10_functional_descriptions

```{r eval=FALSE}
download.file('https://www.arabidopsis.org/download_files/Genes/TAIR10_genome_release/TAIR10_functional_descriptions', 'data/TAIR10_functional_descriptions')
```
```{r}
functions <- read_tsv('data/TAIR10_functional_descriptions')
```
I also downloaded the gene symbols:

```{r eval=FALSE}
esearch -db gene -query "Arabidopsis thaliana [ORGN]" | esummary | xtract -pattern DocumentSummary -element Name,OtherAliases |  awk -F "\t|," '{OFS="\t"}{print $2,$1}' > arabidopsis_gene_symbols.txt
```

```{r}
symbols <- read_tsv('./data/arabidopsis_gene_symbols.txt', col_names = c('Gene','Symbol'))
```

OK now I need to flip the Orthofinder table to then join the functional table

```{r}
ara_groups <- groups %>% select(Orthogroup, `A. thaliana`) %>% filter(!is.na(`A. thaliana`)) %>% separate_rows(Orthogroup, `A. thaliana`, convert=T)
ara_groups <- ara_groups %>% separate(`A. thaliana`, c('Gene','Number'), remove=FALSE)
ara_groups <- left_join(ara_groups, symbols)
ara_groups$Number <- NULL
head(ara_groups)
```

```{r}
ara_joined <- left_join(ara_groups, functions, by=c(`A. thaliana`='Model_name'))
```

You know what? we shouldn't directly compare a dicot (A. thaliana) with monocots (seagrasses), should also add another regular monocot (rice)

```{r}
posi_groups <- groups %>% select(Orthogroup, `P. australis`) %>% filter(!is.na(`P. australis`)) %>% separate_rows(Orthogroup, `P. australis`, convert=T, sep = ', ')
amphi_groups <- groups %>% select(Orthogroup, `A. antarctica`) %>% filter(!is.na(`A. antarctica`)) %>% separate_rows(Orthogroup, `A. antarctica`, convert=T, sep = ', ')
zmar_groups <- groups %>%  select(Orthogroup, `Z. marina`) %>% filter(!is.na(`Z. marina`)) %>% separate_rows(Orthogroup, `Z. marina`, convert=T, sep = ', ')
zmuel_groups <- groups %>%  select(Orthogroup, `Z. muelleri`) %>% filter(!is.na(`Z. muelleri`)) %>% separate_rows(Orthogroup, `Z. muelleri`, convert=T, sep = ', ')
w_australis_groups <- groups %>%  select(Orthogroup, `W. australis`) %>% filter(!is.na(`W. australis`)) %>% separate_rows(Orthogroup, `W. australis`, convert=T, sep = ', ')
s_polyrhiza_groups <- groups %>%  select(Orthogroup, `S. polyrhiza`) %>% filter(!is.na(`S. polyrhiza`)) %>% separate_rows(Orthogroup, `S. polyrhiza`, convert=T, sep = ', ')
#l_gibba_groups <- groups %>%  select(Orthogroup, `L. gibba`) %>% filter(!is.na(`L. gibba`)) %>% separate_rows(Orthogroup, `L. gibba`, convert=T, sep = ', ')
rice_groups<- groups %>%  select(Orthogroup, `O. sativa`) %>% filter(!is.na(`O. sativa`)) %>% separate_rows(Orthogroup, `O. sativa`, convert=T, sep = ', ')
```

```{r}
big_joined <- ara_joined %>% mutate(
  present_in_zmuel = case_when(Orthogroup %in% unique(zmuel_groups$Orthogroup) ~ T,
                               TRUE ~ F),
  present_in_zmar = case_when(Orthogroup %in% unique(zmar_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_amphi = case_when(Orthogroup %in% unique(amphi_groups$Orthogroup) ~ T,
                               TRUE ~ F),
  present_in_posi = case_when(Orthogroup %in% unique(posi_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_waustralis = case_when(Orthogroup %in% unique(w_australis_groups$Orthogroup) ~ T,
                              TRUE ~ F),
  present_in_spolyrhiza = case_when(Orthogroup %in% unique(s_polyrhiza_groups$Orthogroup) ~ T,
                              TRUE ~ F),
 # present_in_lgibba = case_when(Orthogroup %in% unique(l_gibba_groups$Orthogroup) ~ T,
#                              TRUE ~ F),
  present_in_rice = case_when(Orthogroup %in% unique(rice_groups$Orthogroup) ~ T,
                              TRUE ~ F)
  )
```

Done :) Now we have a big table with all *A. thaliana* gene functions and whether these genes are present in clusters of the four seagrasses and three duckweeds.

Let's also add counts - how often is this gene lost?
```{r}
big_joined <- big_joined %>% mutate(Lost_in_Seagrasses = 4 - (present_in_posi + present_in_zmuel + present_in_zmar + present_in_amphi),
                      #Lost_in_duckweeds=3 - (present_in_lgibba + present_in_spolyrhiza + present_in_waustralis),
                      Lost_in_duckweeds=2 - (present_in_spolyrhiza + present_in_waustralis),

                      Lost_in_both = (Lost_in_Seagrasses + Lost_in_duckweeds))
```

```{r}
big_joined %>% writexl::write_xlsx('data/arabidopsis_gene_level_comparison.xlsx')
big_joined %>% filter_at(vars(starts_with('present')), any_vars(. == FALSE)) %>% writexl::write_xlsx('data/arabidopsis_gene_level_comparison_only_losts.xlsx')
```

That was about loss - can we also check for gene family extension using the *A. thaliana* genes?

# Gene family extension

First, let's count how many members each of these orthogroups has per species.
```{r}
ara_count <- ara_joined %>% count(Orthogroup)
posi_count <- posi_groups %>% count(Orthogroup)
amphi_count <- amphi_groups %>% count(Orthogroup)
zmar_count <- zmar_groups %>% count(Orthogroup)
zmuel_count <- zmuel_groups %>% count(Orthogroup)
#lgibba_count <- l_gibba_groups %>% count(Orthogroup)
s_polyrhi_count <- s_polyrhiza_groups %>% count(Orthogroup)
w_aus_count <- w_australis_groups %>% count(Orthogroup)
rice_count <- rice_groups %>% count(Orthogroup)
```

```{r}
counts <- plyr::join_all(list(ara_count, posi_count, amphi_count, zmar_count, zmuel_count, 
                              #lgibba_count, 
                              s_polyrhi_count, w_aus_count, rice_count), by='Orthogroup', type='left') 
names(counts) <- c('Orthogroup', 'A. thaliana', 'P. australis', 'A. antarctica', 'Z. marina', 'Z. muelleri', 
                   #'L. gibba', 
                   'S. polyrhiza', 'W. australis', 'O. sativa')
head(counts)
```

```{r}
joined_counts <- left_join(big_joined, counts, by='Orthogroup') %>% select(-starts_with('present')) %>% 
  select(-starts_with('Lost_i'))
joined_counts %>% writexl::write_xlsx('data/arabidopsis_gene_level_counts.xlsx')
```

# FINAL version redoing - terrestrial vs duckweeds vs algae vs seagrasses

Plotting all species separately is messy as I have so many species. I'm joining the species into three groups: aquatics, terrestrials, and seagrasses!

```{r}
per_group_spec <-
  per_spec %>% mutate(group = case_when(
    name %in% c(
      'A. trichopada',
      'B. distachyon',
      'O. sativa',
      'P. patens',
      'P. trichocarpa',
      'S. moellendorffii',
      'A. thaliana',
      'T. parvula',
      'V. vinifera',
      'Z. mays'
    ) ~ 'Terrestrials',
    name %in% c(
      #'L. gibba',
      'S. polyrhiza',
      'W. australis'
    ) ~ 'Duckweeds',
    name %in% c(
      'C. reinhardtii',
      'O. lucimarinus'
    ) ~ 'Algae',
    name %in% c(
      'P. australis',
      'Z. muelleri',
      'Z. marina',
      'A. antarctica'
      ) ~ 'Seagrasses'
  )) 
```

```{r}
# now I want the data in this format:
# listInput <- list(one = c(1, 2, 3, 5, 7, 8, 11, 12, 13), two = c(1, 2, 4, 5, 
#   10), three = c(1, 5, 6, 7, 8, 9, 10, 12, 13))
groupx <- per_group_spec %>% 
  select(group, Orthogroup) %>% # turn the table around
  deframe() # convert to named vector
```

```{r}
mylistgroup <- lapply(split(groupx, names(groupx)), unname) # yuck - ugly code to convert the named vector to a list
mylistgroup <- lapply(mylistgroup, unique)
```

```{r}
xgroup <- upset(fromList(mylistgroup), order.by='freq', nsets = length(groups) - 1)
```

```{r fig.width=10}
xgroup
```

MUCH better. The biggest chunk of orthogroups is only in terrestrials with 13562 orthogroups. 3845 orthogroups are in duckweeds, seagrasses, terrestrials, but not algae - 2217 orthogroups are unique in seagrasses.

OK now we need to do the GOenrichment for those four groups.

Let's also make an Venn plot, we have only four groups:
```{r fig.width=8}

# code from https://github.com/gaospecial/ggVennDiagram/blob/4cb2aa13c7beae469f9b9836c5d4f94610bd872e/R/ggVennDiagram.R to customise
venn <- Venn(mylistgroup)
data <- process_data(venn)
region_label <- data@region %>%
      dplyr::filter(.data$component == "region") %>%
      dplyr::mutate(percent = paste(round(.data$count*100/sum(.data$count),
                                          digits = 0),"%", sep="")) %>%
      dplyr::mutate(both = paste(.data$count,paste0("(",.data$percent,")"),sep = "\n"))

group_venn <- ggplot() +
    geom_sf(aes_string(fill="count"), data = data@region) +
    geom_sf(aes_string(color = "id"), data = data@setEdge, show.legend = F,
            lty = 'solid', size = 1, color='gray') +
    geom_sf_text(aes_string(label = "name"), data = data@setLabel,
                 size = NA, 
                 color = 'black') +
    theme_void() +
     geom_sf_label(aes_string(label='both'),
                             data = region_label,
                             alpha= 0.0,
                             color = 'black',
                             size = NA,
                             lineheight = 0.85,
                             label.size = NA) +
  scale_fill_gradientn(colours=wes_palette("Zissou1", 100, type = "continuous")) + 
  theme(legend.position = "none") + 
  scale_x_continuous(expand = expansion(mult = .2)) #trick from https://github.com/gaospecial/ggVennDiagram/blob/9435aa0ab4abb470c670ecb938c71576461ccedc/vignettes/using-ggVennDiagram.Rmd#L65

group_venn
```

Nice!

# Write out the per-group GO-terms

Now we pull out all gene-IDs that are only in terrestrials, only in seagrasses, only in aquatics, and unique for each of these three groups, AND we pull out the IDs only present in the four seagrasses each compared with the others.

```{r}
gene_per_spec <- groups %>% pivot_longer(-Orthogroup) %>% 
  filter(!is.na(value)) %>% 
  mutate(group = case_when(
    name %in% c(
      'A. trichopada',
      'B. distachyon',
      'O. sativa',
      'P. patens',
      'P. trichocarpa',
      'S. moellendorffii',
      'A. thaliana',
      'T. parvula',
      'V. vinifera',
      'Z. mays'
    ) ~ 'Terrestrials',
    name %in% c(
     # 'L. gibba',
      'S. polyrhiza',
      'W. australis'
    ) ~ 'Duckweeds',
    name %in% c(
      'C. reinhardtii',
      'O. lucimarinus'
    ) ~ 'Algae',
    name %in% c(
      'P. australis',
      'Z. muelleri', 'Z. marina', 'A. antarctica') ~ 'Seagrasses'
  )) 
```

mylistgroup now looks like this:
key is Aquatics, Seagrasses, Terrestrials, values are OG00000, OG00001, OG00002 etc.

```{r}
present_algae_only <- setdiff(setdiff(setdiff(mylistgroup$Algae, mylistgroup$Duckweeds), 
                         mylistgroup$Seagrasses), mylistgroup$Terrestrials)
present_seagrasses_only <- setdiff(setdiff(setdiff(mylistgroup$Seagrasses, mylistgroup$Duckweeds),                            mylistgroup$Terrestrials), mylistgroup$Algae)
present_terrestrials_only <- setdiff(setdiff(setdiff(mylistgroup$Terrestrials, mylistgroup$Duckweeds), mylistgroup$Seagrasses), mylistgroup$Algae)
present_duckweeds_only <- setdiff(setdiff(setdiff(mylistgroup$Duckweeds, mylistgroup$Terrestrials), mylistgroup$Seagrasses), mylistgroup$Algae)

print(cbind(length(present_algae_only), length(present_seagrasses_only), length(present_terrestrials_only), length(present_duckweeds_only)))

```
OK those numbers fit with the above Venn diagram. I now have four lists of orthogroup names.

I also want the orthogroups present in aquatics + seagrasses, NOT terrestrials

```{r}
present_seagrass_and_aquatics_only <- setdiff(intersect(intersect(mylistgroup$Duckweeds, mylistgroup$Seagrasses), mylistgroup$Algae),
        mylistgroup$Terrestrials)
print(length(present_seagrass_and_aquatics_only))
```
That's the tiny intersection of genes shared between aquatics and seagrasses, let's also get the union of seagrass genes, aquatic genes, and their intersection

```{r}
present_seagrass_and_aquatics_union <- union(mylistgroup$Algae, union(mylistgroup$Duckweeds, mylistgroup$Seagrasses))
print(length(present_seagrass_and_aquatics_union))
```

OK let's also get the intersection in the middle

```{r}
present_all <- intersect(intersect(mylistgroup$Algae, intersect(mylistgroup$Terrestrials, mylistgroup$Duckweeds)), mylistgroup$Seagrasses)
length(present_all)
```

Now we pull out these gene IDs and write them out for files to continue in [GOenrichment](GOenrichment.html).

```{r}
# gene_per_spec has the genes per orthogroup, we need to join with the various set-results we've made
genes_in_all <- gene_per_spec %>% filter(Orthogroup %in% present_all) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
genes_only_in_aquatics_and_seagrasses <- gene_per_spec %>% filter(Orthogroup %in% present_seagrass_and_aquatics_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

genes_present_seagrass_and_aquatics_union <- gene_per_spec %>% filter(Orthogroup %in% present_seagrass_and_aquatics_union) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

genes_only_in_duckweeds <- gene_per_spec %>% filter(Orthogroup %in% present_duckweeds_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
genes_only_in_algae <- gene_per_spec %>% filter(Orthogroup %in% present_algae_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
genes_only_in_seagrasses <- gene_per_spec %>% filter(Orthogroup %in% present_seagrasses_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
genes_only_in_terrestrials <- gene_per_spec %>% filter(Orthogroup %in% present_terrestrials_only) %>% separate_rows(value, sep=',') %>% pull(value) %>% str_trim()
```

Now we have a bunch of lists, we can write them into files for the [GO enrichment part](GOenrichment.html)

```{r}
fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Terrestrials.txt')
writeLines(genes_only_in_terrestrials, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Seagrasses.txt')
writeLines(genes_only_in_seagrasses, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Duckweeds.txt')
writeLines(genes_only_in_duckweeds, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Algae.txt')
writeLines(genes_only_in_algae, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses.txt')
writeLines(genes_only_in_aquatics_and_seagrasses, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_union_of_Seagrass_and_Aquatics_union.txt')
writeLines(genes_present_seagrass_and_aquatics_union, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Aquatics_and_Seagrasses_and_Terrestrials.txt')
writeLines(genes_in_all, fileConn)
close(fileConn)
```

## Dig into seagrass specific differences

In this section, I compare the four seagrasses with each other only, so we can ignore the previous section and go back to the original orthogroups.

```{r}
seagrass_groups <- groups %>% select(Orthogroup, `A. antarctica`, `P. australis`, `Z. marina`, `Z. muelleri`) %>%
  rowwise() %>% 
  mutate(na_count = sum(is.na(c(`A. antarctica`, `P. australis`, `Z. marina`, `Z. muelleri`)))) %>% 
  filter(na_count != 4) %>% 
  select(-na_count)
```

Let's make a Venn diagram like above. We need a list, with the species as keys, and the c(rownumber)

```{r}
big_list <- list()
big_list[['A. antarctica']] <- seagrass_groups %>% select(Orthogroup, `A. antarctica`) %>% filter(!is.na(`A. antarctica`)) %>% pull(Orthogroup)
big_list[['P. australis']] <- seagrass_groups %>% select(Orthogroup, `P. australis`) %>% filter(!is.na(`P. australis`)) %>% pull(Orthogroup)
big_list[['Z. marina']] <- seagrass_groups %>% select(Orthogroup, `Z. marina`) %>% filter(!is.na(`Z. marina`)) %>% pull(Orthogroup)
big_list[['Z. muelleri']] <- seagrass_groups %>% select(Orthogroup, `Z. muelleri`) %>% filter(!is.na(`Z. muelleri`)) %>% pull(Orthogroup)

```

```{r fig.width=8}
venn <- Venn(big_list)
data <- process_data(venn)
# code from https://github.com/gaospecial/ggVennDiagram/blob/4cb2aa13c7beae469f9b9836c5d4f94610bd872e/R/ggVennDiagram.R to customise
region_label <- data@region %>%
      dplyr::filter(.data$component == "region") %>%
      dplyr::mutate(percent = paste(round(.data$count*100/sum(.data$count),
                                          digits = 0),"%", sep="")) %>%
      dplyr::mutate(both = paste(.data$count,paste0("(",.data$percent,")"),sep = "\n"))

seagrass_venn <- ggplot() +
    geom_sf(aes_string(fill="count"), data = data@region) +
    geom_sf(aes_string(color = "id"), data = data@setEdge, show.legend = F,
            lty = 'solid', size = 1, color='gray') +
    geom_sf_text(aes_string(label = "name"), data = data@setLabel,
                 size = NA, fontface='italic',
                 color = 'black') +
    theme_void() +
     geom_sf_label(aes_string(label='both'),
                             data = region_label,
                             alpha= 0,
                             color = 'black',
                             size = NA,
                             lineheight = 0.85,
                             label.size = NA) +
  scale_fill_gradientn(colours=wes_palette("Zissou1", 3, type = "continuous")) + 
  theme(legend.position = "none") + 
  scale_x_continuous(expand = expansion(mult = .2)) #trick from https://github.com/gaospecial/ggVennDiagram/blob/9435aa0ab4abb470c670ecb938c71576461ccedc/vignettes/using-ggVennDiagram.Rmd#L65
seagrass_venn
```

# Both Venns together

Let's save both Venns so I can combine them with other plots in other Rmds.

```{r}
save(group_venn, file='output/group_venn_image.Rdata')
save(seagrass_venn, file='output/seagrass_venn_image.Rdata')
```

```{r fig.height=10, fig.width=8}
group_venn / seagrass_venn +  plot_annotation(tag_levels = 'A')
```

OK now as above, we dig out the gene names.

```{r}
# gene_per_spec has the genes per orthogroup, we need to join with the various set-results we've made
groups_only_in_A_antarctica <- setdiff(setdiff(setdiff(big_list[['A. antarctica']], big_list[['P. australis']]), big_list[['Z. marina']]), big_list[['Z. muelleri']])
genes_only_in_A_antarctica <- gene_per_spec %>% 
  filter(name == 'A. antarctica', Orthogroup %in% groups_only_in_A_antarctica) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

groups_only_in_P_australis <- setdiff(setdiff(setdiff(big_list[['P. australis']], big_list[['A. antarctica']]), big_list[['Z. marina']]), big_list[['Z. muelleri']])
genes_only_in_P_australis <- gene_per_spec %>% 
  filter(name == 'P. australis', Orthogroup %in% groups_only_in_P_australis) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

groups_only_in_Z_marina <- setdiff(setdiff(setdiff(big_list[['Z. marina']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. muelleri']])
genes_only_in_Z_marina <- gene_per_spec %>% 
  filter(name == 'Z. marina', Orthogroup %in% groups_only_in_Z_marina) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

groups_only_in_Z_muelleri <- setdiff(setdiff(setdiff(big_list[['Z. muelleri']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. marina']])
genes_only_in_Z_muelleri <- gene_per_spec %>% 
  filter(name == 'Z. muelleri', Orthogroup %in% groups_only_in_Z_muelleri) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

groups_in_all_seagrasses_union <- union(union(union(big_list[['Z. muelleri']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. marina']])
genes_in_all_seagrasses_union <- gene_per_spec %>% 
  filter(name %in% c('A. antarctica', 'P. australis', 'Z. muelleri', 'Z. marina'), Orthogroup %in% groups_in_all_seagrasses_union) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

groups_in_all_seagrasses_intersect <- intersect(intersect(intersect(big_list[['Z. muelleri']], big_list[['A. antarctica']]), big_list[['P. australis']]), big_list[['Z. marina']])
genes_in_all_seagrasses_intersect <- gene_per_spec %>% 
  filter(name %in% c('A. antarctica', 'P. australis', 'Z. muelleri', 'Z. marina'), Orthogroup %in% groups_in_all_seagrasses_intersect) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

groups_lost_in_A_antarctica <- setdiff(groups_in_all_seagrasses_union, big_list[['A. antarctica']])
genes_lost_in_A_antarctica <- gene_per_spec %>% 
  filter(name != 'A. antarctica', Orthogroup %in% groups_lost_in_A_antarctica) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

groups_lost_in_P_australis <- setdiff(groups_in_all_seagrasses_union, big_list[['P. australis']])
genes_lost_in_P_australis <- gene_per_spec %>% 
  filter(name != 'P. australis', Orthogroup %in% groups_lost_in_P_australis) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

groups_lost_in_Z_marina <- setdiff(groups_in_all_seagrasses_union, big_list[['Z. marina']])
genes_lost_in_Z_marina <- gene_per_spec %>% 
  filter(name != 'Z. marina', Orthogroup %in% groups_lost_in_Z_marina) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

groups_lost_in_Z_muelleri <- setdiff(groups_in_all_seagrasses_union, big_list[['Z. muelleri']])
genes_lost_in_Z_muelleri <- gene_per_spec %>% 
  filter(name != 'Z. muelleri', Orthogroup %in% groups_lost_in_Z_muelleri) %>% 
  separate_rows(value, sep=',') %>% pull(value) %>% str_trim()

```

Now we have a bunch of lists, we can write them into files for the [GO enrichment part](GOenrichment.html)

```{r}
fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_A_antarctica_not_other_seagrasses.txt')
writeLines(genes_only_in_A_antarctica, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_P_australis_not_other_seagrasses.txt')
writeLines(genes_only_in_P_australis, fileConn)
close(fileConn)

fileConn <- file('data/Lost_present_gene_lists/Genes_in_all_seagrasses_union.txt')
writeLines(genes_in_all_seagrasses_union, fileConn)
close(fileConn)

fileConn <- file('data/Lost_present_gene_lists/Genes_in_all_seagrasses_intersect.txt')
writeLines(genes_in_all_seagrasses_intersect, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Z_marina_not_other_seagrasses.txt')
writeLines(genes_only_in_Z_marina, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_only_in_Z_muelleri_not_other_seagrasses.txt')
writeLines(genes_only_in_Z_muelleri, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_lost_in_A_antarctica_not_other_seagrasses.txt')
writeLines(genes_lost_in_A_antarctica, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_lost_in_P_australis_not_other_seagrasses.txt')
writeLines(genes_lost_in_P_australis, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_lost_in_Z_marina_not_other_seagrasses.txt')
writeLines(genes_lost_in_Z_marina, fileConn)
close(fileConn)

fileConn<-file('data/Lost_present_gene_lists/Genes_lost_in_Z_muelleri_not_other_seagrasses.txt')
writeLines(genes_lost_in_Z_muelleri, fileConn)
close(fileConn)
```