### 2023-11-20 analysis/metagenome.Rmd
add metagenomics files. add renv.

---
title: "Visualising metagenome differences"
author: "Philipp Bayer"
date: "20th Nov 2023"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(patchwork)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

Here we compare metagenomes (taxonomy and genes) across Amphibolis vs the others.