### 2023-11-20 analysis/metagenome.Rmd
Expand metagenome analysis

---
title: "Visualising metagenome differences"
author: "Philipp Bayer"
date: "20th Nov 2023"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(UpSetR)
library(microshades)
library(ggtext)
library(patchwork)
library(scales)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

```{r}
trunc_str <- function(latin_name) {
  tmp <- str_split(latin_name, pattern = ' ')[[1]]
  return(paste0(substring(tmp[1], 1, 1), '. ', tmp[2]))
}

```

Here we compare metagenomes (taxonomy and genes) across Amphibolis vs the others.

```{r}
metadata <- read_tsv('./data/metagenome/samples.tsv')
metadata <- metadata |> rename(Sample_species = Species,
                               Sample_replicate = Replicate) |> 
  mutate(Tissue = str_to_title(Tissue))
```

# Taxonomies

We have two taxonomies: one based on CAT, one based on GTDB. GTDB is less complete but more accurate. Let's see what they look like.

# GTDB
```{r}
tax <- read_tsv('./data/metagenome/all_gtdb.summary.tsv.gz')
```

```{r}
clean_tax <- tax |> 
  separate(user_genome, into = c('sample', 'MAG'), sep='\\.') |> 
  separate(classification, into = c('domain', 'phylum', 
                                    'class', 'order', 
                                    'family', 'genus', 
                                    'species'), sep=';') |> 
  mutate(sample = str_replace_all(sample, 'MEGAHIT-', ''), 
         sample = str_replace_all(sample, '_L2', ''))

clean_tax <- clean_tax |> left_join(metadata, by = c('sample'='Sample'))
```

Next, we keep only Posidonia australis and Amphibolis antarctica samples

```{r}
amph_pos_tax <- clean_tax |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
```

## GTDB by order

```{r}
colors <-c(microshades_palette("micro_blue", 4, lightest = FALSE), 
           microshades_palette("micro_purple", 4, lightest = FALSE),
           microshades_palette("micro_green", 4, lightest = FALSE),
          microshades_palette("micro_orange", 4, lightest = FALSE))

```

```{r fig.height=6}
plot_orders_gtdb <- function(amph_pos_tax, group_remainder = TRUE) {
  temp <- amph_pos_tax |> 
    group_by(sample, Tissue, Sample_species) |> 
    mutate(Sample_species = trunc_str(Sample_species)) |> 
    mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
    mutate(order = str_replace_all(order, 'o__','')) |> 
    count(order) 
  if(group_remainder) {
    temp <- temp |> 
    mutate(counted_orders = case_when(n == 1 ~ 'Remainder',
                                      TRUE ~ order))
  } else {
      temp <- temp |> 
        mutate(counted_orders = order)
  }
  temp <- temp |> 
    select(-order) |> 
    group_by(sample, Tissue, Sample_species, n, counted_orders) |> 
    summarise(n = sum(n)) |> 
    filter(counted_orders != '' )

  uniques <- c(setdiff(unique(temp$counted_orders), c('Remainder')), 'Remainder')
  temp |> 
    mutate(counted_orders = factor(counted_orders, levels=uniques)) |> 
    mutate(sample = str_remove(sample, "[A-Z]+")) |> 
    mutate(sample = str_replace(sample, '31', '3')) |> 
    ggplot(aes(x = factor(sample, levels=rev(unique(sample))), y = n, fill=counted_orders)) +
    geom_col() + 
    theme_minimal() + 
    facet_wrap(~Tissue+Sample_species, scales = 'free_y', ncol=2) + 
    coord_flip() + 
    ylab('Number of contigs') + xlab('Replicate') +
    labs(fill = 'Order') +
    scale_y_continuous(breaks=pretty_breaks())  +
    # move legend position to bottom
    theme(legend.position = "bottom",
          strip.text.x = element_markdown()) +
    scale_fill_manual(values = colors)
}
plot_orders_gtdb(amph_pos_tax)
```
Let's also make an Upset plot of that

```{r}
tmp <- amph_pos_tax |> 
  group_by(sample) |> 
  mutate(order = str_replace_all(order, 'o__','')) |> 
  count(order) |> 
  filter(order != '' ) |>
  filter ( n > 1) |> 
  summarise(named_vec = list(order)) %>%
  deframe()
tmp
```
```{r}
upset(fromList(tmp))
```

## GTDB by family

```{r}
amph_pos_tax |> 
  group_by(sample, Tissue) |> 
  mutate(family = str_replace_all(family, 'f__','')) |> 
  count(family) |> 
  filter ( n > 1) |> 
  filter(family != '' ) |> 
  ggplot(aes(x = sample, y = n, fill=family)) +
  geom_col() + 
  theme_minimal() + 
  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  facet_wrap(~Tissue) + coord_flip() + 
  ylab('Number of contigs') + xlab('Sample')
```
```{r}
amph_pos_tax |> group_by(sample, species) |> filter(species != 's__') |> 
  count(species)
```

# CAT
Let's do the same using CAT

```{r}
tax_cat <- read_tsv('./data/metagenome/all_cat.summary.reformatted.tsv.gz')
```

```{r}
clean_tax_cat <- tax_cat |> 
  separate(`# bin`, into = c('sample', 'MAG', 'fa'), sep='\\.') |> 
  mutate(sample = str_replace_all(sample, 'MEGAHIT-', ''), 
         sample = str_replace_all(sample, '_L2', '')) |> 
  select(-fa)

clean_tax_cat <- clean_tax_cat |> left_join(metadata, by = c('sample'='Sample'))
```

Next, we keep only Posidonia australis and Amphibolis antarctica samples

```{r}
amph_pos_tax_cat <- clean_tax_cat |> filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))
```

## CAT by order

```{r fig.height=6}
plot_orders_cat <-
  function(amph_pos_tax_cat, group_remainder = TRUE) {
    temp <- amph_pos_tax_cat |>
      separate(Order, into = c('Order', 'Order_score')) |>
      rename(order = Order) |>
      group_by(sample, Tissue, Sample_species) |>
      mutate(Sample_species = trunc_str(Sample_species)) |>
      mutate(Sample_species = paste0('*', Sample_species, '*')) |>
      count(order)
    if (group_remainder) {
      temp <- temp |>
        mutate(counted_orders = case_when(n == 1 ~ 'Remainder',
                                          TRUE ~ order))
    } else {
      temp <- temp |>
        mutate(counted_orders = order)
    }
    temp <- temp |>
      select(-order) |>
      group_by(sample, Tissue, Sample_species, n, counted_orders) |>
      summarise(n = sum(n)) |>
      filter(counted_orders != '')

    uniques <-
      c(setdiff(unique(temp$counted_orders), c('Remainder')), 'Remainder')
    temp |>
      mutate(counted_orders = factor(counted_orders, levels = uniques)) |>
      mutate(sample = str_remove(sample, "[A-Z]+")) |>
      mutate(sample = str_replace(sample, '31', '3')) |>
      ggplot(aes(
        x = factor(sample, levels = rev(unique(sample))),
        y = n,
        fill = counted_orders
      )) +
      geom_col() +
      theme_minimal() +
      facet_wrap( ~ Sample_species + Tissue, ncol = 2, scales = 'free_y') +
      coord_flip() +
      ylab('Number of contigs') + xlab('Replicate') +
      labs(fill = 'Order') +
      scale_y_continuous(breaks = pretty_breaks())  +
      # move legend position to bottom
      theme(legend.position = "bottom",
            strip.text.x = element_markdown()) +
      scale_fill_manual(values = colors)
  }
plot_orders_cat(amph_pos_tax_cat)
```

## CAT by family

```{r}
amph_pos_tax_cat |> 
  separate(Family, into = c('Family', 'Family_score')) |> 
  group_by(sample, Tissue) |> 
  count(Family) |> 
  filter ( n > 1) |> 
  filter(Family != '' ) |> 
  ggplot(aes(x = sample, y = n, fill=Family)) +
  geom_col() + 
  theme_minimal() + 
  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  facet_wrap(~Tissue) + coord_flip() + 
  ylab('Number of contigs') + xlab('Sample')
```

```{r}
amph_pos_tax_cat |> separate(Species, into = c('Species', 'Species_score'), sep =':') |> filter(!is.na(Species)) |>  group_by(sample, Species) |> 
  count(Species)
```
```{r}
amph_pos_tax_cat |> separate(Species, into = c('Species', 'Species_score'), sep =':') |> filter(!is.na(Species)) |>  group_by(sample, Species) |> 
  count(Species) |> filter(str_detect(Species, 'Acidihalobacter'))
```

There we go :) Which MAGs are those?
```{r}
amph_pos_tax_cat |> filter(str_detect(Species, 'Acidihalobacter'))
```

```{r}
amph_pos_tax_cat |> filter(str_detect(Family, 'Ectothiorhodospiraceae'))
```
Are these in the GTDB classification?
```{r}
amph_pos_tax |> filter(str_detect(family, 'Ectothiorhodospiraceae'))
```
Nope!

# Gene comparison

Let's compare gene presence/absence by names first

```{r}
genes <- read_tsv('././data/metagenome/no_hypothetical.all_genes.tsv.gz')

genes <- genes |> 
  separate(filename, into = c('sample', 'MAG'), sep='\\.', extra = 'drop') |>  # get rid of extra filenaming stuff
  mutate(sample = str_replace(sample, 'MEGAHIT-', '')) |> 
  separate(MAG, into = c('MAG', 'rest'), sep ='/') |> 
  select(-rest)
```

```{r}
genessplit <- genes |> separate(gene, into = c('gene', 'gene_copy_number'), sep='_')
```

```{r}
genessplit <- genessplit |> mutate(sample = str_replace(sample, '_L2', '')) |> 
  left_join(metadata, by = c('sample'='Sample')) |> 
  filter(Sample_species %in% c('Amphibolis antarctica', 'Posidonia australis'))

```

```{r}
count_diffs <- genessplit |> 
  group_by(Sample_species, gene) |> 
  count(gene) |> 
  pivot_wider(names_from = Sample_species, values_from = n) |> 
  mutate(difference = `Amphibolis antarctica` - `Posidonia australis`) |> 
  arrange(desc(difference))
```

```{r}
count_diffs |> filter(`Posidonia australis` <100)
```

We hypothesise that the ACC precursor-existing genes in Amphibolis lead to different microbiomes. ACC is broken down by ACC deaminases, which are present in Amphibolis but not Posidonia. Let's see if we can find ACC deaminases in the metagenomes (*acdS*).

```{r}
genessplit |> filter(gene == 'acdS') |> 
  group_by(Sample_species, Tissue, Sample_replicate) |> 
  count(gene) |> 
  mutate(Sample_species = trunc_str(Sample_species)) |> 
  mutate(Sample_species = paste0('*', Sample_species, '*')) |> 
  ggplot(aes(x= interaction(factor(Sample_replicate, levels=4:1), Tissue), y = n, fill = Sample_species)) + 
  geom_col() + 
  facet_wrap(~Sample_species) +
  coord_flip()+ theme_minimal() + 
  scale_y_continuous(breaks=pretty_breaks()) +
  ylab('Total number of *acdS*-containing MAGs') +
  xlab('Sample species') + 
  theme(legend.position = 'none',
        strip.text.x =  element_markdown(),
        axis.title.x = element_markdown())
```

What are those acdS-containing MAGs?

```{r}
genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG'), multiple = 'any')
```
Lots of unknown taxonomies :(

```{r}
genessplit |> filter(gene == 'acdS') |> left_join(amph_pos_tax, by=c('sample', 'MAG'), multiple = 'any') |> 
  select(-c(Tissue.y, Sample_species.y)) |>
  rename(Tissue = Tissue.x, Sample_species = Sample_species.x) |> 
  plot_orders_gtdb(group_remainder = FALSE)
  #select(sample, MAG, domain:species)
```

```{r}
genessplit |> filter(gene == 'acdS') |> 
  left_join(amph_pos_tax_cat, by=c('sample', 'MAG')) |> 
    select(-c(Tissue.y, Sample_species.y)) |>
  rename(Tissue = Tissue.x, Sample_species = Sample_species.x) |> 
  plot_orders_cat(group_remainder = FALSE) #
  #select(sample, MAG, Superkingdom:Species) |>

```