### 2023-11-22 analysis/index.Rmd

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

This website collects comparisons of *Amphibolis* and *Posidonia* comparative analyses.

# Results

[Phylogenies](OTT.html) - plotting just phylogenies

[Clustering results](Orthofinder.html) - analysing Orthofinder clustering results

[R gene plotting](plotRgenes.html) - making phylogenies and showing R-gene number differences

[Gene Loss comparison by reads](GeneLossComparisonReads.html) - making heatmaps and barcharts to compare read alignments

[GO enrichment](GOenrichment.html) - running topGO and REVIGO to make nice plots of GO-enrichment

[GO and Ortho](GO_and_ortho.html) - merging plots from GO enrichment and Orthofinder analysis

[Metagenomics](metagenome.html) - visualising metagenome differences across Posidonia/Amphibolis

# Cutting floor

Additional analyses that are not in the paper.

[Repeat comparison plots](plot.html) - Comparing genome sizes and repeat/TE content

[MSA visualisation](MSA.html) - visualising MSAs and phylogenies for gene clusters of interest (*SOS3*, *EIN3*)

[3D structure](ThreeDStructure.html) - visualising the predicted 3D structure of *SOS3*/*EIN3*